### 2014-02-26 cloudpaper/paper.tex
added ScienceCloud paper to repo

% THIS IS SIGPROC-SP.TEX - VERSION 3.1
% WORKS WITH V3.2SP OF ACM_PROC_ARTICLE-SP.CLS
% APRIL 2009
%
% It is an example file showing how to use the 'acm_proc_article-sp.cls' V3.2SP
% LaTeX2e document class file for Conference Proceedings submissions.
% ----------------------------------------------------------------------------------------------------------------
% This .tex file (and associated .cls V3.2SP) *DOES NOT* produce:
%       1) The Permission Statement
%       2) The Conference (location) Info information
%       3) The Copyright Line with ACM data
%       4) Page numbering
% ---------------------------------------------------------------------------------------------------------------
% It is an example which *does* use the .bib file (from which the .bbl file
% is produced).
% REMEMBER HOWEVER: After having produced the .bbl file,
% and prior to final submission,
% you need to 'insert'  your .bbl file into your source .tex file so as to provide
% ONE 'self-contained' source file.
%
% Questions regarding SIGS should be sent to
% Adrienne Griscti ---> griscti@acm.org
%
% Questions/suggestions regarding the guidelines, .tex and .cls files, etc. to
% Gerald Murray ---> murray@hq.acm.org
%
% For tracking purposes - this is V3.1SP - APRIL 2009

\documentclass{acm_proc_article-sp}

\usepackage{url} % for url citation
\usepackage{epstopdf}

\begin{document}

\title{Towards Non-Consumptive Cloud Computing for Digital Humanities}

\author{}

\maketitle
\begin{abstract}
As the amount of digitized archives grows, compute technology plays an increasing important role in digital humanities research. The advent of cloud computing offers researcher computation capabilities to process tens of thousands of archives which goes beyond their desktop machines can process. However, for copyrighted archives, it poses security challenges to the compute platform. In this paper, we investigate non-consumptive research, a research that allows user access copyrighted archives while prevents archives reconstruction during user access, in cloud computing. We build an infrastructure as a service (IaaS) cloud framework that makes use of data capsule to enforce non-consumptive research policy in virtual machine. This paper presents our architecture design and implementation details, and outlines next steps for future development.
\end{abstract}

% A category with the (minimum) three required fields
\category{H.3.4}{Systems and Software}{Distributed systems}[cloud computing, data capsule]
%A category including the fourth, optional field follows...
\category{K.6.m}{Miscellaneous}{Security}[non-consumptive research]

\terms{Architecture, Design, Implementation}

\keywords{Cloud computing, Non-consumptive research, data capsule} % NOT required for Proceedings

\section{Introduction}
As the amount of digitized archives keeps growing, researchers of digital humanities are facing with materials that they could ever hope to read in life time. For example, some researchers checked some of Ian Watt's claims against 3,600 novels between 18th and 19th century \cite{Example}. It is tedious, if possible, for human being to do this task by hitting the library stacks, checking out books and examining book contents. Therefore, digitized archives with computer access significantly changes the way digital humanities researchers perform research. Computer technology has been applied to facilitate digital humanities research on massive digitized materials. For instance, researchers run topic modeling algorithm on a set of digitized books to identify common topics among books \cite{LDA}, which cannot be discovered easily by just reading the books.

However, the massive amount of digitized texts also poses challenges on data analysis. First, compute capabilities required vary and depend on the analysis tasks as well as amount of texts. For example, running topic modeling algorithm generally requires more CPU cycles and memory than generating a word cloud based on word frequency. Fixed hardware setting of user desktop is not suitable to deal with such dynamic workloads. Second, high performance computing(HPC) provides a way for user to allocate resource dynamically based on her workload. However, HPC is batch oriented and digital humanities researchers usually need an interactive way to access the compute resources. Thus cloud computing is preferred in such case. Third, some of the digitized contents are copyrighted. Allowing user algorithms run directly on copyrighted texts opens a vulnerable channel that non-consumptive constraint may be violated. According to \cite{BookSettlement}, non-consumptive research is a research that no action or set of actions on the part of users, either acting alone or in cooperation with other users over the duration of one or multiple sessions can result in sufficient information gathered from a collection of copyrighted works to reassemble pages from the collection. In other words, a user may steal copyrighted content while she runs her algorithm. We extend the definition above to include the situation that a user computer is compromised by malicious programs which acts on user\rq s behave to steal copyrighted content.

We propose a non-consumptive cloud framework to address challenges mentioned above. On one hand, cloud computing, which has elastic computing feature, is a natural fit for various analysis tasks. Researcher can create virtual machines shipped with different compute capabilities based on the need. Once the analysis is done, virtual machine is wiped out and resources are released for other users to share. On the other hand, the virtual machine of our framework is equipped with data capsule \cite{Borders:2009:PCD:1855768.1855791} to provide non-consumptive policy to prevent copyrighted leakage under the assumption that virtual machine is compromised. In data capsule, one virtual machine has two modes, i.e. maintenance mode and secure mode. In maintenance mode, user can access network and install software freely, but cannot access copyrighted texts. In secure mode, copyrighted texts become accessible to user while network access and file system access are highly constrained. Users are only allowed to access a predefined set of network addresses and write to a specific volume which is only visible in secure mode. Every other changes user made to the system in secure mode, except for the ones made to special volume, is lost when mode is switched from secure to maintenance. This is guard against the situation that copyrighted texts is saved in VM in secure mode and copied out through network in maintenance.

Our framework is implemented as a sub system within HathiTrust Research Center (HTRC) \cite{HTRC}. HTRC aims at enabling computational access for nonprofit and educational users to perform research on digitized texts. Through various services and tools provided from HTRC, user can access digitized texts, and run text mining algorithms on HTRC compute resources.

This paper is organized as follows. Section \ref{title:related} summaries related work. Section \ref{title:overview} presents an overview of HTRC and data capsule, and identifies the design requirements for our system. Section \ref{title:design} describes details of our system design and implementation. Section \ref{title:future} outlines future work and section \ref{title:conclusion} concludes.

\section{Related Work} \label{title:related}

Researchers have already leveraged the power of cloud computing in digital archives. Teregowda et al. employed cloud computing as underlying infrastructure for CiteSeerX, which is a public search engine and digital library for scientific and academic papers \cite{Teregowda:cloud}. Rosenthal et al. investigated the usage of cloud in their digital preservation system -- LOCKSS for libraries \cite{Rosenthal:preservation}. Our system differs from those systems in the sense that we investigate and promote non-consumptive research in the cloud platform.

Generally speaking, cloud computing can be categories as SaaS (software as a service), PaaS (platform as a service), and IaaS (infrastructure as a service) \cite{Armbrust:EECS-2009-28}. Our cloud framework intends to provide a weaker form of IaaS (infrastructure as a service). That is user can configure some of the VM specs, e.g., operation system, number of CPU, memory, and disk, but not the network, nor the volume. There are numbers of existing cloud platforms support IaaS, either in a commercial way or in an open source way. Amazon EC2 \cite{EC2} is probably the most well-known IaaS platform.  On regard to open source cloud platform, Eucalyptus \cite{Nurmi:2009:EOC:1577849.1577895} is an open source IaaS platform that provides Amazon compatible cloud environment. OpenStack \cite{OpenStack} is another open source IaaS platform that gains increasing popularity. Researchers investigated different cloud platforms \cite{Sempolinski:cloud,vonLaszewski:2012:CMC:2353730.2353779}. The reasons we do not build our system on existing open source platforms are two folds. First, existing cloud platforms introduce numerous complexities for the data capsule which enforces non-consumptive research policy in VM. Those complexities are from unnecessary features for digital humanities researchers. For example, researcher does not need to configure a software defined network to run her text mining algorithms. Second, existing cloud platforms are designed for general purpose and expose many vulnerable channels. For example, allowing user to configure network would pose significant threats to our systems. Therefore, instead of using existing cloud platform, we build a cloud framework around data capsule, which satisfies digital humanities researchers\rq need and non-consumptive research requirement. Our research also illuminates the path that data capsule could potentially be applied to more general cloud platforms.

Related work of Data Capsule.

\section{Overview} \label{title:overview}
This section presents background of HathiTrust Research Center (HTRC), data capsule and identifies threat model as well as design requirements for our system.

\subsection{HathiTrust Research Center Background} \label{subtitle:background}
HTRC consists of number of services that facilitate nonprofit and educational users to access and research on texts from HathiTrust Digital Library. Figure \ref{fig:htrc} below shows an overview of HTRC from an interaction point of view. HTRC provides users with a complexity hiding interface, left bar on Figure \ref{fig:htrc}, which shadows HTRC service complexity. The central cloud shape represents the entire HTRC system which can be divided into 4 categories: 1) metadata and data subsystem which provide metadata, raw text, and statistical information access; 2) data mining subsystem that embraces several tools for users to mine corpus; 3) job execution subsystem that runs various text mining algorithms on HPC environment; 4) auxiliary services subsystem that supports other services, e.g., user authentication service.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.3]{figures/htrc}
  \caption{HathiTrust Research Center Infrastructure.}
  \label{fig:htrc}
\end{figure}

\subsection{Threat Model}

The system is designed to allow user to access copyrighted data in a compromised operating system which runs in a virtual machine (VM). Neither the operating system, the driver, nor the applications (except for user’s application running against HTRC Data service) are trusted. Although our system allows arbitrary behaviors in VM, it does require some trusted components. The virtual machine monitor (VMM) that runs virtual machines is part of the trusted computing base (TCB). The hosts where VMM locates are also part of the TCB. The internal network among VMs is part of the TCB, i.e., firewall policy is sufficient to prevent attack from one VM to another one. User’s application that runs against HTRC Data service is considered trustworthy, although it could potentially encode the copyrighted data and deliver it as final result. We require user to sign a legal agreement before she uses our system. We intend to add some result analysis to automatically detect data leakage and leave this as future work.

\subsection{Data Capsule}

The data capsule was developed by Borders et al. \cite{Borders:2009:PCD:1855768.1855791}. It enforces two modes on a single virtual machine, i.e., maintenance mode and secure mode, to provide secure access to sensitive data. In maintenance mode, user is free to access internet and install software in VM. While in secure mode, sensitive data stored in a secure volume will be attached to the VM and become visible to user. The network access is blocked in secure mode to prevent data leakage by network. When user switches from secure mode to maintenance mode, changes made to places other than volume are wiped out so that sensitive data cannot leak through copying to local disk neither. Only changes made to volume is persistent. Therefore, in data capsule, user can access sensitive data and leakage cannot take place neither through network channel nor through disk.

Data capsule was mainly developed for general usage in single machine. We adapt and extend data capsule based on the context of HTRC. Sensitive data in HTRC is copyrighted text which is exposed through a API in Restful web service instead of staying in volume. Data capsule needs to enable the network channel between web service serving data and virtual machine. On regard to future HTRC services, data capsule should have a flexible network control mechanism. Besides, to carry out analysis tasks, researchers need persistent storage to store temporal results for iterative analysis purpose. Thus, the role of secure volume in data capsule is changed from mainly serving read to mainly serving write. In addition, data capsule is designed to work in single machine. To provide a cloud environment across multiple machines to multiple researcher, we need another layer on top of data capsule to coordinate different machines. We propose a web service layer, which is responsible for management effects e.g., resource allocation, request scheduling, status maintenance, and etc, on top of the data capsule to implement the cloud environment. Section \ref{title:design} gives details of design and implementation of data capsule in HTRC context.

\subsection{Design Requirements} \label{subtitle:requirement}
Our non-consumptive cloud framework falls into the second category mentioned in section \ref{subtitle:background} through provisioning a secure environment for user to conduct research against our data, probably copyrighted. We target digital humanities researchers and aim at providing a user friendly environment for user to access copyrighted text while preventing copyrighted text leakage. We identify following design requirements.
\begin{enumerate}
  \item It provides a cloud environment that makes better use of our backend resources;
  \item It prevents texts copied out from our system through network, and interference among users;
  \item It audits users activities;
  \item It is easy for researchers to use. Researchers have the privilege to configure the software and hardware environment to meet their needs;
  \item It can support large scale data process paradigm, e.g., MapReduce.
\end{enumerate}

\section{Design and Implementation} \label{title:design}
This section presents the details of the architecture and implementation of our system.

\subsection{Architecture}

Figure \ref{fig:architecture} delineates our architecture design. It consists of three layers, physical layer, web service layer and web front end layer from bottom to top. The physical layer includes physical machines and hypervisor software that run virtual machines. It also includes the data capsule implementation which is scripts that wrap hypervisor commands to perform multiple tasks. Database, image store and volume store also sit in this layer. Database is used by the web service to maintain persistent states for virtual machines and different operations. Virtual machine images as well as secure volumes locate in NFS, a share file system. The web service layer is the central controller of the system. It is responsible for resource allocation, request scheduling, state maintenance, and failover on physical layer. It also has an audit component to log user activities. Finally we expose the system through the web font end layer. It consists of a web UI where user interacts with our system and a user authentication server which validates user identification. To validate user identity, we use OAuth 2.0 \cite{OAuth2} for user authentication.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.33]{figures/layers}
  \caption{Overall architecture design.}
  \label{fig:architecture}
\end{figure}

\subsection{Workflows}
This section presents control flows within our system components.

\subsubsection{Virtual machine operations} \label{subsubtitle:control}
Figure \ref{fig:control} describes the control flow in our system. First user logs into web UI, which delegates the authentication to OAuth 2 server which validates user login credentials. Details of user authentication is presented in section \ref{subsubtitle:auth}. Upon validation success, user can manipulate the VM through the web UI. We use VM creation request as an example here. The UI forwards the request along with an authentication token to web service. When web service receives the request, it validates the token against OAuth 2 server again. The reason for double validation is web UI might be compromised and web service needs to validate requests coming from web UI. If validation successes, web service will prepare to invoke corresponding hypervisor script. It makes scheduling decision on which host to launch VM, allocates ports for VM, retrieves image information from database, and etc. Once the preparation is done, it will call hypervisor script remotely to launch a VM on a particular host, monitor the response and update VM state in database. Meanwhile web service will return user the information required to log into VM, e.g., VNC port. Other VM operations (shutdown, delete, launch, switch mode from one another) follow the same path described above.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.35]{figures/workflow-control}
  \caption{System control flow.}
  \label{fig:control}
\end{figure}

\subsubsection{Authentication} \label{subsubtitle:auth}
We use OAuth 2.0 as our authentication protocol \cite{OAuth2}. A dedicated server with OAuth 2.0 implementation is setup to authenticate user. Following steps summarize the authentication procedure as presented in Figure \ref{fig:auth}.
\begin{enumerate}
  \item User requests login to the web UI;
  \item Web UI redirects user to the authentication server which requires user login credentials;
  \item Authentication server sends a token back to web UI to indicate validation successes;
  \item After receiving the token, web UI can forward user requests to web service along with the token;
  \item Web service layer would make sure that the request is valid for the given user through validating token against authentication server. Web service layer would also use the token to retrieve the user's identity from authentication server.
\end{enumerate}

\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.5]{figures/workflow-authentication}
  \caption{Authentication flow.}
  \label{fig:auth}
\end{figure}

\subsubsection{Virtual machine access}

Section \ref{subsubtitle:control} presents an overview of different VM operations. This section focuses on the mechanism we provide to user to access VM. There are two possibilities that copyrighted data may leak from our system during user access. One of them is through network. For one VM, we intend to block arbitrary network access when a user accesses copyrighted data, while keeping network open for the user to download and configure software environment when she does not access copyrighted data. The data capsule provides two VM mode, i.e. maintenance mode and secure mode, to support different network control for the same VM. In maintenance mode, user can access network without any constrains, although the HTRC data service is blocked by firewall policy. In secure mode, user can only access HTRC data service. Any other network accesses are denied. Figure \ref{fig:vmaccess} gives an example. First user creates and starts a VM through the web UI as described in section \ref{subsubtitle:control}. After the VM is up and running, the user can log in to it through any VNC client with proper information provided from web UI, e.g., host name and port number. Note that VNC session is the only channel that user can access VM in secure mode. We block SSH traffic, which is a typical way to access VM in most cloud platforms e.g., Amazon EC2, OpenStack, because it allows user to copy data out by scp command. By default, user logs into VM in maintenance mode. User can install software from internet, upload programs from her desktop, and etc. If user wants to access HTRC Data service, she needs to switch VM mode from maintenance to secure through web UI. In practice, VNC screen will be frozen for a short time (usually 2 to 5 seconds) during the mode switch. In secure mode, user does not have network access except for HTRC Data service.

The other possibility of copyrighted data leakage is to copy data from secure mode to maintenance mode in local disk and leak through network in maintenance mode. In secure mode, copyrighted data may be copied from HTRC Data service in local disk. When VM is switched to maintenance mode, the copyrighted data in local disk is visible in maintenance mode and can be copied out through network. To prevent such leakage, data capsule checkpoints VM image before it goes to secure mode. That is to take a snapshot of the VM by using qemu command. When VM switches from secure mode to maintenance mode, data capsule restores the checkpoint, i.e, the snapshot image, through qemu command. Therefore, all the data written to VM local disk in secure mode will be wiped out when VM switches from secure mode to maintenance mode. Meanwhile, to allow user to persist her data in secure mode, data capsule provides a secure volume, which is only visible in secure mode and shown as an external drive to the VM, to let user write data to. The secure volume stores data persistently and will be detached from VM when VM is in maintenance mode.

A typical workflow for a user to access the VM is depicted in figure \ref{fig:vmaccess}. First, the user creates and starts the VM through web UI. Second, user logs in to VM in maintenance mode through VNC client (step 1). Third, user configures software setting through internet (step 2). Forth, user switches VM mode from maintenance to secure, accesses HTRC Data service and writes data to secure volume (step 3, 4).

Final result release??

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.45]{figures/workflow-HTRCData}
  \caption{Virtual machine access flow.}
  \label{fig:vmaccess}
\end{figure}

\subsubsection{Final results review}

Under construction.

In summary, our architecture fulfills the design requirements in section \ref{subtitle:requirement}. First, the web front end and web service use OAuth 2.0 to authenticate user which denies unauthorized users. Second, web service coordinates a number of machines equipped with data capsule to provide a cloud environment and log all user activities. Through the log files, we are able to tell who did what. Third, the data capsule blocks network channels when VM is in secure mode so that no copy out can take place. Forth, researcher can manage and manipulate her virtual machines through a simple web UI which is more user friendly than a command line tool. Besides, researcher can log in to virtual machine through VNC which provisions GUI desktop instead of terminal. In addition, researcher is allowed to configure the hardware setting when creating the VM from web UI and software setting in VM in maintenance mode to meet her needs. These effects help lower the usage burdens for digital humanities researchers. Last but not least, the web service can setup a Hadoop cluster for researcher to run MapReduce jobs although more efforts are needed to configure data capsule properly in a MapReduce context. We leave it as future work.

\subsection{Implementation Details}

\subsubsection{Physical layer implementation}
Under construction.

\subsubsection{Web service layer implementation}
The web service is implemented as an asynchronous Restful service in Jersey framework []. Since operations in hypervisor layer usually take some time to finish, we decouple the response and actual execution to provide good responsiveness. For instance, when a start VM request arrives, web service returns \emph{start-pending} as VM state immediately rather than waits until VM startup finishes and returns \emph{start} state. The web service also takes into account failover on hypervisor layer. It automatically retries connection failure and timeouts long running ssh session. The web service is implemented in a multi-threaded way so that multiple operations on different VM can happen simultaneously.

\section{Future Work} \label{title:future}
We are in the process of implementing final results review mechanism. Once that part is in place, we expect to release our system to humanities researchers for use and get feedbacks from them. There are numerous next steps we can take. First, we intend to build MapReduce cluster automatically with data capsule in place to expand the utilization of data capsule. Second, we are interested in providing researchers with extractive features that comply with non-consumptive research policy. Potential qualified features include word frequency, term document matrix, etc. Third, we will investigate the approach of adding random noises to final result so that user is unable to reconstruct original texts while the accuracy of final results is bounded. For example, after adding noises, the lower bound of accuracy is 90\% which may be sufficient enough to researchers.

\section{Conclusion} \label{title:conclusion}

This paper proposes a cloud framework which complies with non-consumptive research policy to facilitate digital humanities research. The framework is built on top of data capsule and data capsule is extending to adapt to a cloud environment.  We present design details of the framework along with implementation details. Most parts of the system are implemented. We are in the process of implementing the final result review mechanism and plan to release our system to digital humanities researchers in the next couple of months.

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}

%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns

% That's all folks!
\end{document}

### 2014-02-26 cloudpaper/paper.tex
Updates to paper

@@
 \begin{abstract}
-As the amount of digitized archives grows, compute technology plays an increasing important role in digital humanities research.
-The advent of cloud computing offers researcher computation capabilities to process tens of thousands of archives which goes beyond their desktop machines can process.
-However, for copyrighted archives, it poses security challenges to the compute platform.
-In this paper, we investigate non-consumptive research, a research that allows user access copyrighted archives while prevents archives reconstruction during user access, in cloud computing.
-We build an infrastructure as a service (IaaS) cloud framework that makes use of data capsule to enforce non-consumptive research policy in virtual machine.
-This paper presents our architecture design and implementation details, and outlines next steps for future development.
+As the number of digitized archives grows, computing technology is playing an
+increasingly important role in digital humanities research. The advent of cloud
+computing offers researchers the ability to quickly process tens of thousands
+of archives, going far beyond the capabilities of their desktop machines.
+However, copyrighted archives pose a crucial security challenge to this
+otherwise promising computational research platform. In this paper, we
+investigate \emph{non-consumptive research}, that is, research that allows
+users access to copyrighted archives via cloud computing, while preventing the
+release of the archive data during user access. We have built an infrastructure
+as a service (IaaS) cloud framework that makes use of storage capsules to
+enforce a non-consumptive research policy within a virtual machine.  This paper
+presents our architecture, design, and implementation details, and outlines the
+next steps for future development.
 \end{abstract}
@@
 % A category with the (minimum) three required fields
-\category{H.3.4}{Systems and Software}{Distributed systems}[cloud computing, data capsule]
+\category{H.3.4}{Systems and Software}{Distributed Systems}[Cloud Computing, Storage Capsules]
 %A category including the fourth, optional field follows...
-\category{K.6.m}{Miscellaneous}{Security}[non-consumptive research]
+\category{K.6.m}{Miscellaneous}{Security}[Non-consumptive Research]
 
@@
 
-\keywords{Cloud computing, Non-consumptive research, data capsule} % NOT required for Proceedings
+\keywords{Cloud computing, Non-consumptive Research, Storage Capsules} % NOT required for Proceedings
 
@@
 
-Related work of Data Capsule.
+Storage Capsules \cite{Borders:2009:PCD:1855768.1855791} is a mechanism for
+containing sensitive data within a virtualized environment with the goal of
+minimizing the available channels to leak that data.  Although Capsules
+provides no guarantees on the integrity of the stored data -- which could be
+modified or deleted by an attacker -- this poses little issue when Capsules is
+applied to non-consumptive research, where the copyrighted source data can
+simply be made read-only.  Capsules uses virtual machine snapshotting and
+simple policy-based restriction mechanisms to allow a user to enter a
+restricted ``secure'' mode, where network and other data channels are
+restricted in exchange for access to the data being protected.  After using the
+protected data, when the user returns to normal use of the system, all changes
+to the system except those made to the designated secure data are forgotten,
+and the system returns to the state it was in before the switch to secure mode.
+This use model is readily applicable to non-consumptive research, in that a
+researcher can administer their system (e.g. install any required software
+tools) and then switch to ``secure'' mode to perform their analysis.
 
@@
 
+\subsection{Data Capsule}
+
+Storage Capsules was developed by Borders et al.
+\cite{Borders:2009:PCD:1855768.1855791}. It enforces two modes on a single
+virtual machine, i.e., maintenance mode and secure mode, to provide secure
+access to sensitive data. In maintenance mode, user is free to access internet
+and install software in VM. While in secure mode, sensitive data stored in a
+secure volume will be attached to the VM and become visible to user. The
+network access is blocked in secure mode to prevent data leakage by network.
+When user switches from secure mode to maintenance mode, changes made to places
+other than volume are wiped out so that sensitive data cannot leak through
+copying to local disk neither. Only changes made to volume is persistent.
+Therefore, in data capsule, user can access sensitive data and leakage cannot
+take place neither through network channel nor through disk.
+
+Data capsule was mainly developed for general usage in single machine. We adapt
+and extend data capsule based on the context of HTRC. Sensitive data in HTRC is
+copyrighted text which is exposed through a API in Restful web service instead
+of staying in volume. Data capsule needs to enable the network channel between
+web service serving data and virtual machine. On regard to future HTRC
+services, data capsule should have a flexible network control mechanism.
+Besides, to carry out analysis tasks, researchers need persistent storage to
+store temporal results for iterative analysis purpose. Thus, the role of secure
+volume in data capsule is changed from mainly serving read to mainly serving
+write. In addition, data capsule is designed to work in single machine. To
+provide a cloud environment across multiple machines to multiple researcher, we
+need another layer on top of data capsule to coordinate different machines. We
+propose a web service layer, which is responsible for management effects e.g.,
+resource allocation, request scheduling, status maintenance, and etc, on top of
+the data capsule to implement the cloud environment. Section \ref{title:design}
+gives details of design and implementation of data capsule in HTRC context.
+
 \subsection{Threat Model}
 
-The system is designed to allow user to access copyrighted data in a compromised operating system which runs in a virtual machine (VM).
-Neither the operating system, the driver, nor the applications (except for user’s application running against HTRC Data service) are trusted.
-Although our system allows arbitrary behaviors in VM, it does require some trusted components.
-The virtual machine monitor (VMM) that runs virtual machines is part of the trusted computing base (TCB).
-The hosts where VMM locates are also part of the TCB.
-The internal network among VMs is part of the TCB, i.e., firewall policy is sufficient to prevent attack from one VM to another one.
-User’s application that runs against HTRC Data service is considered trustworthy, although it could potentially encode the copyrighted data and deliver it as final result.
-We require user to sign a legal agreement before she uses our system.
-We intend to add some result analysis to automatically detect data leakage and leave this as future work.
-
-\subsection{Data Capsule}
-
-The data capsule was developed by Borders et al.
-\cite{Borders:2009:PCD:1855768.1855791}.
-It enforces two modes on a single virtual machine, i.e., maintenance mode and secure mode, to provide secure access to sensitive data.
-In maintenance mode, user is free to access internet and install software in VM.
-While in secure mode, sensitive data stored in a secure volume will be attached to the VM and become visible to user.
-The network access is blocked in secure mode to prevent data leakage by network.
-When user switches from secure mode to maintenance mode, changes made to places other than volume are wiped out so that sensitive data cannot leak through copying to local disk neither.
-Only changes made to volume is persistent.
-Therefore, in data capsule, user can access sensitive data and leakage cannot take place neither through network channel nor through disk.
-
-Data capsule was mainly developed for general usage in single machine.
-We adapt and extend data capsule based on the context of HTRC.
-Sensitive data in HTRC is copyrighted text which is exposed through a API in Restful web service instead of staying in volume.
-Data capsule needs to enable the network channel between web service serving data and virtual machine.
-On regard to future HTRC services, data capsule should have a flexible network control mechanism.
-Besides, to carry out analysis tasks, researchers need persistent storage to store temporal results for iterative analysis purpose.
-Thus, the role of secure volume in data capsule is changed from mainly serving read to mainly serving write.
-In addition, data capsule is designed to work in single machine.
-To provide a cloud environment across multiple machines to multiple researcher, we need another layer on top of data capsule to coordinate different machines.
-We propose a web service layer, which is responsible for management effects e.g., resource allocation, request scheduling, status maintenance, and etc, on top of the data capsule to implement the cloud environment.
-Section \ref{title:design} gives details of design and implementation of data capsule in HTRC context.
+The system is designed to allow users to access copyrighted data in a potentially
+compromised operating system which runs in a virtual machine (VM). Neither the
+operating system, the driver, nor the applications (except for user’s
+application running against HTRC Data service) are trusted. Although our system
+allows arbitrary behaviors in VM, it does require some trusted components. The
+virtual machine monitor (VMM) that runs virtual machines is part of the trusted
+computing base (TCB). The hosts where VMM locates are also part of the TCB. The
+internal network among VMs is part of the TCB, i.e., firewall policy is
+sufficient to prevent attack from one VM to another one. User’s application
+that runs against HTRC Data service is considered trustworthy, although it
+could potentially encode the copyrighted data and deliver it as final result.
+We require user to sign a legal agreement before she uses our system. We intend
+to add some result analysis to automatically detect data leakage and leave this
+as future work.
 

### 2014-02-27 cloudpaper/paper.tex
added my text and incorporated version 3 of the paper

--- 2014-02-26 cloudpaper/paper.tex
@@
 
-\title{Towards Non-Consumptive Cloud Computing for Digital Humanities}
+\title{Cloud Computing Data Capsule for Non-consumptive Text Analysis}
 
-\author{}
+\numberofauthors{8}
+\author{
+\alignauthor
+Jiaan Zeng \\
+       \affaddr{School of Informatics and Computing Indiana University}\\
+       \email{jiaazeng@indiana.edu}
+\alignauthor
+Guangchen Ruan \\
+       \affaddr{School of Informatics and Computing Indiana University}\\
+       \email{gruan@indiana.edu}
+\alignauthor 
+Alex Crowell \\
+       \affaddr{Department of Computer Science and Engineering University of Michigan}\\
+       \email{crowella@umich.edu}
+\and  % use '\and' if you need 'another row' of author names
+\alignauthor 
+Atul Prakash \\
+       \affaddr{Department of Computer Science and Engineering University of Michigan}\\
+       \email{aprakash@umich.edu}
+% 5th. author
+\alignauthor Beth Plale \\
+       \affaddr{School of Informatics and Computing Indiana University}\\
+       \email{plale@indiana.edu}
+}
 
@@
 \begin{abstract}
-As the amount of digitized archives grows, compute technology plays an increasing important role in digital humanities research.
-The advent of cloud computing offers researcher computation capabilities to process tens of thousands of archives which goes beyond their desktop machines can process.
-However, for copyrighted archives, it poses security challenges to the compute platform.
-In this paper, we investigate non-consumptive research, a research that allows user access copyrighted archives while prevents archives reconstruction during user access, in cloud computing.
-We build an infrastructure as a service (IaaS) cloud framework that makes use of data capsule to enforce non-consumptive research policy in virtual machine.
-This paper presents our architecture design and implementation details, and outlines next steps for future development.
+As digital archives grow in size, the interest by humanities and informatics
+researchers alike increases in computational investigation of a corpus.  The
+cloud computing model of ``data as a service'' which co-locating computation with
+a data collection of value, allows for scales of computational analysis that
+far exceeds what a desktop machine can provide. However, when data collections
+are protected, such as digitized books are through copyright, there are
+additional challenges to provisioning an analytical framework. We investigate
+issues in text analysis on a protected corpus in a cloud environment that
+prevents reconstruction of part or whole of the corpus.  The proposed framework
+makes use of a novel notion of "data capsule" to enforce non-consumptive
+research policy using virtual machines.
 \end{abstract}
@@
 
-\keywords{Cloud computing, Non-consumptive research, data capsule} % NOT required for Proceedings
+\keywords{Cloud computing, Large-scale Text Mining, Non-consumptive Research, Storage Capsules} % NOT required for Proceedings
 
 \section{Introduction}
-As the amount of digitized archives keeps growing, researchers of digital humanities are facing with materials that they could ever hope to read in life time.
-For example, some researchers checked some of Ian Watt's claims against 3,600 novels between 18th and 19th century \cite{Example}.
-It is tedious, if possible, for human being to do this task by hitting the library stacks, checking out books and examining book contents.
-Therefore, digitized archives with computer access significantly changes the way digital humanities researchers perform research.
-Computer technology has been applied to facilitate digital humanities research on massive digitized materials.
-For instance, researchers run topic modeling algorithm on a set of digitized books to identify common topics among books \cite{LDA}, which cannot be discovered easily by just reading the books.
+As increasing amounts of digitized text become available from sources such as university libraries, researchers encounter access to more material than they could ever hope to read in life time, creating opportunities for new forms of research.
+For example, research has checked literary critic Ian Watts' claims against 3,600 novels written between the 18th and 19th centuries \cite{Example}.
+While this tedious task could be carried out by hand, digitized texts coupled with text mining tools, enables new discoveries.
+One can apply topic modeling such as Latent Dirichlet allocation (LDA)~\cite{LDA}) at the book level to narrow from millions of texts to a few thousand, then page level LDA to zero down to relevant pages, effectively expanding the reach of research investigation of the informatics or digital humanities researcher.
 
-However, the massive amount of digitized texts also poses challenges on data analysis.
-First, compute capabilities required vary and depend on the analysis tasks as well as amount of texts.
-For example, running topic modeling algorithm generally requires more CPU cycles and memory than generating a word cloud based on word frequency.
-Fixed hardware setting of user desktop is not suitable to deal with such dynamic workloads.
-Second, high performance computing(HPC) provides a way for user to allocate resource dynamically based on her workload.
-However, HPC is batch oriented and digital humanities researchers usually need an interactive way to access the compute resources.
-Thus cloud computing is preferred in such case.
-Third, some of the digitized contents are copyrighted.
-Allowing user algorithms run directly on copyrighted texts opens a vulnerable channel that non-consumptive constraint may be violated.
-According to \cite{BookSettlement}, non-consumptive research is a research that no action or set of actions on the part of users, either acting alone or in cooperation with other users over the duration of one or multiple sessions can result in sufficient information gathered from a collection of copyrighted works to reassemble pages from the collection.
-In other words, a user may steal copyrighted content while she runs her algorithm.
-We extend the definition above to include the situation that a user computer is compromised by malicious programs which acts on user\rq s behave to steal copyrighted content.
+Texts digitized from research libraries have restrictions on their use; most obviously copyright for content that is post about 1923.
+While most in the legal community would agree that text mining qualifies as fair use, the text mining cannot as a byproduct leak entire books to web pages.
+This kind of limited analytical use on the data we call \textit{non-consumptive research}.
+Non-consumptive research we define as \textit{No action or set of actions on the part of users, either
+acting alone or in cooperation with other users over the duration of one or multiple sessions can result in sufficient information gathered from a collection of copyrighted works to reassemble pages from the collection.}
 
-We propose a non-consumptive cloud framework to address challenges mentioned above.
-On one hand, cloud computing, which has elastic computing feature, is a natural fit for various analysis tasks.
-Researcher can create virtual machines shipped with different compute capabilities based on the need.
-Once the analysis is done, virtual machine is wiped out and resources are released for other users to share.
-On the other hand, the virtual machine of our framework is equipped with data capsule \cite{Borders:2009:PCD:1855768.1855791} to provide non-consumptive policy to prevent copyrighted leakage under the assumption that virtual machine is compromised.
-In data capsule, one virtual machine has two modes, i.e.
-maintenance mode and secure mode.
-In maintenance mode, user can access network and install software freely, but cannot access copyrighted texts.
-In secure mode, copyrighted texts become accessible to user while network access and file system access are highly constrained.
-Users are only allowed to access a predefined set of network addresses and write to a specific volume which is only visible in secure mode.
+In this paper we propose a novel solution to non-consumptive research that uses a data capsule approach~\cite{Borders:2009:PCD:1855768.1855791} preventing misuse.
+The work is being carried out in the context of the HathiTrust Research Center~\cite{HTRC}, a set of software and services that currently provide text analysis capability over public domain books.
+HTRC is in the process of securing access to 12 million volumes of text data and images from research libraries across the country.
+
+Text analytics over a large corpus of sensitive data poses challenges for several reasons.
+First, text mining tools vary in their computational demand.
+Analysis could take the form of simple access to a database of features; could involve a linear traversal through digitized text; or could be a computationally intensive natural language processing task like LDA.
+Second, while high performance computing (HPC) gives a user resources dynamically based on workload, HPC systems remain largely batch oriented which is out of phase with digital humanities researcher requiring high levels of interaction with texts.
+The cloud computing data-as-service model avoids the problem of limited interactivity.
+Third, and the biggest challenge of the three, is that digitized texts can have restrictions on their access.
+Allowing interactive access on texts user algorithms run directly on copyrighted texts opens a vulnerable channel that non-consumptive constraint may be violated.
+
+We propose a non-consumptive cloud framework for data analytics on sensitive corpora.
+Using a remote virtual machine (VM) model, researchers can build a VM configured with software and tooling based on their need.
+Once an analysis is finished, the VM is wiped out and resources released for other users to share.
+VMs offer no inherent protection, however.
+Our approach extends the virtual machine by turning it into a \textit{data capsule}\cite{Borders:2009:PCD:1855768.1855791} that prevents leakage of copyrighted content in the event that the VM is compromised or data analysis routines malicious.
+The HTRC Data Capsule provides the virtual machine with two modes: a maintenance mode during which a user can access the network and install software freely, but cannot access copyrighted data; and secure mode where copyrighted texts become accessible to user while the network access and file system access is highly constrained.
+In the latter mode, users are allowed only  access to a predefined set of network addresses and write to a specific volume which is only visible in secure mode.
 Every other changes user made to the system in secure mode, except for the ones made to special volume, is lost when mode is switched from secure to maintenance.
-This is guard against the situation that copyrighted texts is saved in VM in secure mode and copied out through network in maintenance.
+This is to guard against the situation that copyrighted texts is saved in VM in secure mode and copied out through network in maintenance.
 
-Our framework is implemented as a sub system within HathiTrust Research Center (HTRC) \cite{HTRC}.
-HTRC aims at enabling computational access for nonprofit and educational users to perform research on digitized texts.
-Through various services and tools provided from HTRC, user can access digitized texts, and run text mining algorithms on HTRC compute resources.
+This paper introduces the architecture and design of the HTRC Data Capsule.  The research question driving HTRC Data Capsule imposes five major requirements as follows:
 
-This paper is organized as follows.
-Section \ref{title:related} summaries related work.
-Section \ref{title:overview} presents an overview of HTRC and data capsule, and identifies the design requirements for our system.
+\begin{itemize}
+\item Non-consumptive research: that is, provides safe handling of large volumes of data, and
+can ensure that the read restrictions of the definition hold
+\item Openness: users are not limited to using a known set of algorithms, and instead are
+expected to experiment with their own algorithms
+\item Efficiency: It is not possible to analyze algorithms for conformance prior to
+execution
+\item Low cost and scale:  Run at large-scale and low cost to users
+\item  Long term and broad value:  framework is designed for adoption for other
+purposes
+\end{itemize}
+
+%Jiaan,  these above are the requirements from the proposal.  These replace the requirements that were in Sec 3, which were much more implementation focused and not appropriate for a research paper.
+
+This paper is organized as follows: Section \ref{title:related} summaries related work and \ref{title:overview} presents an overview of HTRC and data capsule, and identifies the design requirements for our system.
 Section \ref{title:design} describes details of our system design and implementation.
@@
 
-Researchers have already leveraged the power of cloud computing in digital archives.
+Research that leverages the power of cloud computing for text analysis has been done.
 Teregowda et al.
-employed cloud computing as underlying infrastructure for CiteSeerX, which is a public search engine and digital library for scientific and academic papers \cite{Teregowda:cloud}.
+employ cloud computing as underlying infrastructure for CiteSeerX, which is a public search engine and digital library for scientific and academic papers \cite{Teregowda:cloud}.
 Rosenthal et al.
-investigated the usage of cloud in their digital preservation system -- LOCKSS for libraries \cite{Rosenthal:preservation}.
-Our system differs from those systems in the sense that we investigate and promote non-consumptive research in the cloud platform.
+propose LOCKSS for digital preservation; LOCKSS uses a cloud model in their digital preservation system~\cite{Rosenthal:preservation}.
+Our system differs from those systems because of the protection requirements of non-consumptive research in the cloud platform.
 
-Generally speaking, cloud computing can be categories as SaaS (software as a service), PaaS (platform as a service), and IaaS (infrastructure as a service) \cite{Armbrust:EECS-2009-28}.
-Our cloud framework intends to provide a weaker form of IaaS (infrastructure as a service).
-That is user can configure some of the VM specs, e.g., operation system, number of CPU, memory, and disk, but not the network, nor the volume.
-There are numbers of existing cloud platforms support IaaS, either in a commercial way or in an open source way.
-Amazon EC2 \cite{EC2} is probably the most well-known IaaS platform.
-On regard to open source cloud platform, Eucalyptus \cite{Nurmi:2009:EOC:1577849.1577895} is an open source IaaS platform that provides Amazon compatible cloud environment.
-OpenStack \cite{OpenStack} is another open source IaaS platform that gains increasing popularity.
+There are numerous cloud platforms that support IaaS, either commercially or in open source.
+Amazon EC2 \cite{EC2} is probably the most well-known IaaS platform, but there is Eucalyptus \cite{Nurmi:2009:EOC:1577849.1577895}, and OpenStack \cite{OpenStack}.
 Researchers investigated different cloud platforms \cite{Sempolinski:cloud,vonLaszewski:2012:CMC:2353730.2353779}.
-The reasons we do not build our system on existing open source platforms are two folds.
-First, existing cloud platforms introduce numerous complexities for the data capsule which enforces non-consumptive research policy in VM.
-Those complexities are from unnecessary features for digital humanities researchers.
-For example, researcher does not need to configure a software defined network to run her text mining algorithms.
+We do not build our system on existing open source cloud platforms for two reasons: first, existing cloud platforms introduce numerous complexities for the data capsule which has to enforce the non-consumptive research policy in the VM.
+The complexities stem from avoiding imposing unnecessary burdens on the digital humanities researcher.
+For example, the researcher should not need to configure a software defined network to run her text mining algorithms.
 Second, existing cloud platforms are designed for general purpose and expose many vulnerable channels.
@@
 
-Related work of Data Capsule.
+Storage capsules \cite{Borders:2009:PCD:1855768.1855791} is a mechanism for
+containing sensitive data within a virtualized environment with the goal of
+minimizing the available channels to leak that data.  Although storage capsules
+provide no guarantees on the integrity of the stored data -- which could be
+modified or deleted by an attacker -- this poses little issue when storage
+capsules is applied to non-consumptive research, where the copyrighted source
+data can simply be made read-only.  Additionally, the storage capsules design
+is well-suited to non-consumptive research, in that it is able to maintain
+strong security guarantees on the sensitive data while being highly permissive
+to users, who may need to install custom software or perform other actions that
+would require the granting of many privileges in a normal environment.  The
+application of storage capsules we present in this paper currently does not
+provide the same level of protection against covert channels that is provided
+in the original storage capsules system, but improvements in the implementation
+are being made that will eventually provide greater protection than that given
+in the original work.
 
 \section{Overview} \label{title:overview}
-This section presents background of HathiTrust Research Center (HTRC), data capsule and identifies threat model as well as design requirements for our system.
+This section presents background of the HathiTrust Research Center software and services, data capsule and identifies threat model as well as design requirements for our system.
 
 \subsection{HathiTrust Research Center Background} \label{subtitle:background}
-HTRC consists of number of services that facilitate nonprofit and educational users to access and research on texts from HathiTrust Digital Library.
-Figure \ref{fig:htrc} below shows an overview of HTRC from an interaction point of view.
-HTRC provides users with a complexity hiding interface, left bar on Figure \ref{fig:htrc}, which shadows HTRC service complexity.
-The central cloud shape represents the entire HTRC system which can be divided into 4 categories: 1) metadata and data subsystem which provide metadata, raw text, and statistical information access; 2) data mining subsystem that embraces several tools for users to mine corpus; 3) job execution subsystem that runs various text mining algorithms on HPC environment; 4) auxiliary services subsystem that supports other services, e.g., user authentication service.
+The HathiTrust Research Center is a set of software and services to carry out computational analysis using digitized books from the HathiTrust Digital Library for research and educational use.
+Figure \ref{fig:htrc} is a cartoon diagram of the system.
+A researcher accesses the system through one or more front ends (shown on right).
+The system in its simplest form satisfies a user need by constructing for the user a \textit{Workset} representing selected text mining tools, selected subsets of data from either the HathiTrust digital repository, or from feature sets that have been extracted in advance, and data from other sources.
+This Workset bundle is executed on a large-scale compute resource.
+As is shown by the arrow pointing to both compute resource and front end (complexity hiding interface), text analysis is a highly interactive process, thus challenging the use of HPC resources shown, even if, like Big Red 2 at Indiana University, they are architected to handle data-intensive computations.
+The HTRC system is modularly architected using a web services paradigm (i.e., REST interfaces).
+It utilizes a Solr index, Cassandra noSQL store, both of which are sharded across half-dozen machines for higher replicability and availability.
+Analytical tools include the SEASR suite of text mining tools, with ongoing extension of support to R and i-Python.
 
-\begin{figure}[h]
+The HTRC architecture uses OAuth2 authentication and has an auditing service.
+While this architecture is adequate for analysis algorithms that have been vetted by HTRC, it cannot support analysis algorithms that originate with an external researcher.
+This latter issue is the purview of the HTRC Data Capsule.
+
+\begin{figure}[tbh]
   \centering
   \includegraphics[scale=0.3]{figures/htrc}
-  \caption{HathiTrust Research Center Infrastructure.}
+  \caption{Functional diagram of HathiTrust Research Center software and services}
   \label{fig:htrc}
@@
 
+\subsection{Storage Capsules}
+
+Storage capsules is a system developed by Borders et al.
+\cite{Borders:2009:PCD:1855768.1855791} that allows privileged access to
+sensitive data while also restricting the channels through which that data can
+be released.  Capsules uses virtual machine snapshotting and simple
+policy-based restriction mechanisms to allow a user to enter a ``secure'' mode,
+where network and other data channels are restricted in exchange for access to
+the data being protected.  After using the protected data, when the user
+returns to normal use of the system, which we call ``maintenance'' mode, all
+changes to the system except those made to the designated secure data are
+forgotten, and the system returns to the state it was in before the switch to
+secure mode, thereby regaining full Internet access and the ability to make
+persistent changes to the virtual machine, such as installing software.  In
+this way, network and storage channels cannot be used to leak the sensitive
+data.  This usage model is readily applicable to non-consumptive research, in
+that a researcher can administer their system (e.g. install any required
+software tools) and then switch to ``secure'' mode to perform their analysis.
+
+Because storage capsules was originally developed with use on a personal
+machine in mind, it had to be adapted and extended to work within the HTRC
+cloud environment. In HTRC, sensitive data is copyrighted text that is exposed
+through a Restful web service API instead of within a machine. Storage capsules
+manages the network channel between virtual machines and the web services
+serving data to them. In consideration of potential future HTRC services,
+storage capsules has a flexible network control mechanism that allows network
+policies for different modes to be written by administrators.
+
+Additionally, in order to provide a cloud environment across multiple machines
+and to a large number of researchers, there is a need for another layer on top
+of storage capsules that coordinates different machines. We therefore
+implemented a web service layer that is responsible for management of virtual
+machines (e.g., resource allocation, request scheduling, status maintenance)
+above the storage capsules layer.  Section \ref{title:design} gives details of
+the design and implementation of storage capsules and the web service layer in
+the context of HTRC.
+
 \subsection{Threat Model}
 
-The system is designed to allow user to access copyrighted data in a compromised operating system which runs in a virtual machine (VM).
-Neither the operating system, the driver, nor the applications (except for user’s application running against HTRC Data service) are trusted.
-Although our system allows arbitrary behaviors in VM, it does require some trusted components.
-The virtual machine monitor (VMM) that runs virtual machines is part of the trusted computing base (TCB).
-The hosts where VMM locates are also part of the TCB.
-The internal network among VMs is part of the TCB, i.e., firewall policy is sufficient to prevent attack from one VM to another one.
-User’s application that runs against HTRC Data service is considered trustworthy, although it could potentially encode the copyrighted data and deliver it as final result.
-We require user to sign a legal agreement before she uses our system.
-We intend to add some result analysis to automatically detect data leakage and leave this as future work.
+Under our usage model, users access the copyrighted data through remotely
+accessed virtual machines that read the data from a network-accessed data
+service.  The virtual machine presented by the user is not a part of the trusted
+computing base (TCB).  We assume the possibility of malware being installed as
+well as other remotely initiated attacks on the VM.  These attacks could
+potentially compromise the entire operating system and install a rootkit that
+are undetectable to the end user.  The end users themselves are considered to
+act in good faith, but this does not preclude the possibility of them
+unwittingly allowing the system to be compromised.  We believe this is a
+reasonable assumption since users are required to sign a legal agreement before
+using our system.  Beyond the virtual machine itself, the virtual machine
+manager (VMM) and the host it runs on are both trusted, and fall within the TCB.
+This includes system services that enforce network and data access policies for
+the virtual machines.  Finally, the HTRC Data service itself is also a part of
+the TCB.
 
-\subsection{Data Capsule}
+Users have VNC access to their virtual machines so that they have a graphical
+interface to the machine.  However, this access does admittedly provide a
+channel for potential data leak.  For now, we apply our assumption that the end
+user acts in good faith, and also assume that they are the only one accessing
+the virtual machine.  For future work, we can use analysis of the channel as a
+means to automatically detect potential abuses of it for data leak purposes.
+Additionally, a channel is provided to the user for releasing results when their
+research is complete.  Although we currently intend to allow released results to
+be downloaded via a link sent to the end user's email inbox, we note that the
+released data could also be subjected to manual or automated review to detect
+potential abuses.  We also leave this for future work.
 
-The data capsule was developed by Borders et al.
-\cite{Borders:2009:PCD:1855768.1855791}.
-It enforces two modes on a single virtual machine, i.e., maintenance mode and secure mode, to provide secure access to sensitive data.
-In maintenance mode, user is free to access internet and install software in VM.
-While in secure mode, sensitive data stored in a secure volume will be attached to the VM and become visible to user.
-The network access is blocked in secure mode to prevent data leakage by network.
-When user switches from secure mode to maintenance mode, changes made to places other than volume are wiped out so that sensitive data cannot leak through copying to local disk neither.
-Only changes made to volume is persistent.
-Therefore, in data capsule, user can access sensitive data and leakage cannot take place neither through network channel nor through disk.
-
-Data capsule was mainly developed for general usage in single machine.
-We adapt and extend data capsule based on the context of HTRC.
-Sensitive data in HTRC is copyrighted text which is exposed through a API in Restful web service instead of staying in volume.
-Data capsule needs to enable the network channel between web service serving data and virtual machine.
-On regard to future HTRC services, data capsule should have a flexible network control mechanism.
-Besides, to carry out analysis tasks, researchers need persistent storage to store temporal results for iterative analysis purpose.
-Thus, the role of secure volume in data capsule is changed from mainly serving read to mainly serving write.
-In addition, data capsule is designed to work in single machine.
-To provide a cloud environment across multiple machines to multiple researcher, we need another layer on top of data capsule to coordinate different machines.
-We propose a web service layer, which is responsible for management effects e.g., resource allocation, request scheduling, status maintenance, and etc, on top of the data capsule to implement the cloud environment.
-Section \ref{title:design} gives details of design and implementation of data capsule in HTRC context.
-
-\subsection{Design Requirements} \label{subtitle:requirement}
-Our non-consumptive cloud framework falls into the second category mentioned in section \ref{subtitle:background} through provisioning a secure environment for user to conduct research against our data, probably copyrighted.
-We target digital humanities researchers and aim at providing a user friendly environment for user to access copyrighted text while preventing copyrighted text leakage.
-We identify following design requirements.
-\begin{enumerate}
-  \item It provides a cloud environment that makes better use of our backend resources;
-  \item It prevents texts copied out from our system through network, and interference among users;
-  \item It audits users activities;
-  \item It is easy for researchers to use. Researchers have the privilege to configure the software and hardware environment to meet their needs;
-  \item It can support large scale data process paradigm, e.g., MapReduce.
-\end{enumerate}
+Another potential threat is that of covert channels between virtual machines
+that run on the same host machine.  For instance, a virtual machine running in
+secure mode could possibly make use of such covert channels to leak data to a
+coresident virtual machine running in maintenance mode, which can in turn leak
+the data anywhere it pleases.  We acknowledge that this could pose a serious
+threat, but leave addressing it for future work.
 
 \section{Design and Implementation} \label{title:design}
-This section presents the details of the architecture and implementation of our system.
 
-\subsection{Architecture}
-
-Figure \ref{fig:architecture} delineates our architecture design.
-It consists of three layers, physical layer, web service layer and web front end layer from bottom to top.
-The physical layer includes physical machines and hypervisor software that run virtual machines.
+ The high level architecture of the HTRC Data Capsule, see Figure \ref{fig:architecture}, consists of three layers shown bottom to top: a physical layer, a web service layer and web front-end layer.
+The physical layer consists of physical machines and hypervisor software that run the virtual machines of the data capsule.
 It also includes the data capsule implementation which is scripts that wrap hypervisor commands to perform multiple tasks.
 Database, image store and volume store also sit in this layer.
-Database is used by the web service to maintain persistent states for virtual machines and different operations.
-Virtual machine images as well as secure volumes locate in NFS, a share file system.
+The database is used by the web service layer to maintain persistent states for virtual machines and different operations.
+
+Virtual machine images as well as secure volumes locate on NFS, a distributed file system.
 The web service layer is the central controller of the system.
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering
   \includegraphics[scale=0.33]{figures/layers}
-  \caption{Overall architecture design.}
+  \caption{HTRC Data Capsule architecture}
   \label{fig:architecture}
@@
 \subsection{Workflows}
-This section presents control flows within our system components.
+
+There are three major control flows within the HTRC Data Capsule system:   virtual machine creation, authentication, and VM access.  Each are described in turn.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[ht]
   \centering
@@
 
+%Jiaan,  this section needs to tie back around and wrap in the requirements laid out in section 1.
+The implementation requirements in Sec 3 are now gone.
+In addressing the broader goals in Sec 1,  how well does the work so far address these requirements?
+What's been done and what hasn't been addressed (for instance: we haven't tried for generality yet.)  This is still a 'towards' paper so can and should be honest about where we are and where open work still exists
+
 Under construction.
 
-In summary, our architecture fulfills the design requirements in section \ref{subtitle:requirement}.
 First, the web front end and web service use OAuth 2.0 to authenticate user which denies unauthorized users.
@@
 \subsubsection{Physical layer implementation}
-Under construction.
+
+The backend that manages virtual machines is implemented as a set of Linux
+shell scripts that manage the \texttt{qemu} VMM in order to implement the
+functionality of storage capsules.  These scripts are remotely run by the web
+service via a secure shell (SSH) connection.  \texttt{qemu} provides live
+snapshotting functionality that serves an important role in realizing our
+storage capsules implementation.  When a user transitions to secure mode, a
+snapshot is taken of the entire virtual machine state via the backend scripts,
+and any further use of the system is recorded in memory and does not persist
+upon shutdown or any other operation.  Upon return to maintenance mode, the
+snapshot taken earlier is restored, and all system progress that followed it is
+forgotten.
+
+Virtual machines are isolated from the rest of the host's networking through the
+use of a tap interface.  Their connection is then bridged to host networking.
+Network policies for both modes are implemented using iptables rules.  These
+rules can be written in custom policy files that can be specified on mode
+switch, allowing for any arbitrary policy desired by the administrator.  In
+addition to the policy rules, some basic rules are always enforced, for example
+preventing communication between virtual machines residing on the same host.
 
@@
 \section{Acknowledgments}
+This research funded through a grant from the Alfred P. Sloan Foundation, grant \# xxxxxx
+%ask Nicole to give you the grant number; I'm not finding it.   This needs to be the number the Sloan foundation knows, not IU's internal grant #
 

### 2014-02-27 cloudpaper/paper.tex
More minor changes
settled on 'Data Capsules'

@@
 \alignauthor 
-Alex Crowell \\
+Alexander Crowell \\
        \affaddr{Department of Computer Science and Engineering University of Michigan}\\
@@
 % A category with the (minimum) three required fields
-\category{H.3.4}{Systems and Software}{Distributed Systems}[Cloud Computing, Storage Capsules]
+\category{H.3.4}{Systems and Software}{Distributed Systems}[Cloud Computing, Data Capsules]
 %A category including the fourth, optional field follows...
@@
 
-\keywords{Cloud computing, Large-scale Text Mining, Non-consumptive Research, Storage Capsules} % NOT required for Proceedings
+\keywords{Cloud computing, Large-scale Text Mining, Non-consumptive Research, Data Capsules} % NOT required for Proceedings
 
@@
 
-Storage capsules \cite{Borders:2009:PCD:1855768.1855791} is a mechanism for
+Data capsules \cite{Borders:2009:PCD:1855768.1855791} is a mechanism for
 containing sensitive data within a virtualized environment with the goal of
-minimizing the available channels to leak that data.  Although storage capsules
+minimizing the available channels to leak that data.  Although data capsules
 provide no guarantees on the integrity of the stored data -- which could be
-modified or deleted by an attacker -- this poses little issue when storage
+modified or deleted by an attacker -- this poses little issue when data 
 capsules is applied to non-consumptive research, where the copyrighted source
-data can simply be made read-only.  Additionally, the storage capsules design
+data can simply be made read-only.  Additionally, the data capsules design
 is well-suited to non-consumptive research, in that it is able to maintain
@@
 would require the granting of many privileges in a normal environment.  The
-application of storage capsules we present in this paper currently does not
+application of data capsules we present in this paper currently does not
 provide the same level of protection against covert channels that is provided
-in the original storage capsules system, but improvements in the implementation
+in the original data capsules system, but improvements in the implementation
 are being made that will eventually provide greater protection than that given
@@
 
-\subsection{Storage Capsules}
+\subsection{Data Capsules}
 
-Storage capsules is a system developed by Borders et al.
+Data capsules is a system developed by Borders et al.
 \cite{Borders:2009:PCD:1855768.1855791} that allows privileged access to
@@
 
-Because storage capsules was originally developed with use on a personal
+Because data capsules was originally developed with use on a personal
 machine in mind, it had to be adapted and extended to work within the HTRC
 cloud environment. In HTRC, sensitive data is copyrighted text that is exposed
-through a Restful web service API instead of within a machine. Storage capsules
+through a Restful web service API instead of within a machine. Data capsules
 manages the network channel between virtual machines and the web services
 serving data to them. In consideration of potential future HTRC services,
-storage capsules has a flexible network control mechanism that allows network
+data capsules has a flexible network control mechanism that allows network
 policies for different modes to be written by administrators.
@@
 and to a large number of researchers, there is a need for another layer on top
-of storage capsules that coordinates different machines. We therefore
+of data capsules that coordinates different machines. We therefore
 implemented a web service layer that is responsible for management of virtual
 machines (e.g., resource allocation, request scheduling, status maintenance)
-above the storage capsules layer.  Section \ref{title:design} gives details of
-the design and implementation of storage capsules and the web service layer in
+above the data capsules layer.  Section \ref{title:design} gives details of
+the design and implementation of data capsules and the web service layer in
 the context of HTRC.
@@
 shell scripts that manage the \texttt{qemu} VMM in order to implement the
-functionality of storage capsules.  These scripts are remotely run by the web
+functionality of data capsules.  These scripts are remotely run by the web
 service via a secure shell (SSH) connection.  \texttt{qemu} provides live
 snapshotting functionality that serves an important role in realizing our
-storage capsules implementation.  When a user transitions to secure mode, a
+data capsules implementation.  When a user transitions to secure mode, a
 snapshot is taken of the entire virtual machine state via the backend scripts,

### 2014-07-15 ccsw14/paper.tex
Created CCSW'14 paper

\documentclass{acm_proc_article-sp}

\usepackage{epstopdf}

\begin{document}

\title{Securing Remote Access to Sensitive Data}

%\numberofauthors{2} 
%
%\author{
%\alignauthor Alexander Crowell \\
%       \affaddr{Computer Science and Engineering Division}\\
%       \affaddr{University of Mighigan}\\
%       \affaddr{Ann Arbor, MI, USA}\\
%       \email{crowella@umich.edu}
%\alignauthor Atul Prakash \\
%       \affaddr{Computer Science and Engineering Division}\\
%       \affaddr{University of Mighigan}\\
%       \affaddr{Ann Arbor, MI, USA}\\
%       \email{aprakash@umich.edu}
%}

\maketitle
\begin{abstract}

As we enter an era of ``big data", where analysis of large-scale data is
revealing important insights into a variety of fields, there is an ever greater
demand for access to new data wherein the potential for new insights may lie.
However, in many cases, this need for access conflicts with the desire to
protect data privacy.  Indeed, many types of important data present just such a
dilemma, including copyrighted data, personal information of individuals, and
data containing state or corporate secrets.

In this paper, we propose Data Capsules, a system that is designed to enable
more secure access to sensitive data while maintaining reasonable guarantees of
data privacy.  Data Capsules uses virtualization to provide remote users with a
secure, privileged environment in which they can access and analyze sensitive
data using software of their choice, while minimizing the available channels for
data leaks, including covert channels.  Our early implementation realizes most
of this protection, with the exception of the end user's remote desktop
connection.

\end{abstract}

\category{D.4.6}{Operating Systems}{Security and Protection}
\category{H.3.4}{Systems and Software}{Distributed Systems}[cloud computing, data capsules]

\keywords{data capsules, data privacy}

\section{Introduction}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Praesent eget
tristique dolor, sit amet volutpat lacus. Curabitur nec enim id sapien molestie
feugiat\cite{clark:pct}. Integer adipiscing sapien urna, rhoncus vehicula elit pellentesque at.
Aliquam pulvinar feugiat dui quis porta. Cras et erat justo. Phasellus at neque
sed nisl tincidunt consectetur ut eget ante. Donec porta urna ut bibendum
cursus. Aenean euismod pellentesque turpis vitae placerat. Nam orci turpis,
pharetra et lacus sit amet, accumsan semper massa.

%Morbi metus nulla, tristique
%nec arcu a, ornare fermentum lectus. Curabitur interdum libero non massa
%laoreet commodo. Morbi porta, risus sit amet rhoncus egestas, felis diam
%sollicitudin massa, eu iaculis purus sem in turpis. Duis commodo facilisis
%purus non volutpat. Ut ullamcorper metus in ultrices tempus. Integer nec quam
%vitae justo dictum vehicula a non augue. Proin sit amet gravida libero.
%
Donec ornare iaculis nisl, sit amet luctus metus varius ac. Nam eget tincidunt
magna. Sed convallis pulvinar eros sed consequat. Pellentesque posuere orci id
leo molestie, ac rutrum dolor pretium. In orci lectus, sollicitudin nec tellus
non, dictum egestas ante. Ut venenatis libero risus, sed auctor lorem
scelerisque at. Suspendisse porta elementum magna, id ornare erat placerat at.
In tincidunt dolor porta justo rhoncus elementum.

Nullam viverra turpis nisi, id consequat urna pulvinar eu. Praesent ac pulvinar
nisl. Integer vel est a libero molestie mollis id vel tortor. Morbi ante odio,
cursus et volutpat a, pharetra at erat. Pellentesque habitant morbi tristique
senectus et netus et malesuada fames ac turpis egestas. Curabitur dignissim,
purus at auctor laoreet, magna mi auctor nulla, et mollis eros nisl vel lectus.
Integer bibendum purus ac adipiscing bibendum.

Donec ornare iaculis nisl, sit amet luctus metus varius ac. Nam eget tincidunt
magna. Sed convallis pulvinar eros sed consequat. Pellentesque posuere orci id
leo molestie, ac rutrum dolor pretium. In orci lectus, sollicitudin nec tellus
non, dictum egestas ante. Ut venenatis libero risus, sed auctor lorem
scelerisque at. Suspendisse porta elementum magna, id ornare erat placerat at.
In tincidunt dolor porta justo rhoncus elementum.

Nullam viverra turpis nisi, id consequat urna pulvinar eu. Praesent ac pulvinar
nisl. Integer vel est a libero molestie mollis id vel tortor. Morbi ante odio,
cursus et volutpat a, pharetra at erat. Pellentesque habitant morbi tristique
senectus et netus et malesuada fames ac turpis egestas. Curabitur dignissim,
purus at auctor laoreet, magna mi auctor nulla, et mollis eros nisl vel lectus.
Integer bibendum purus ac adipiscing bibendum.

\section{Related Work}
Donec ornare iaculis nisl, sit amet luctus metus varius ac. Nam eget tincidunt
magna. Sed convallis pulvinar eros sed consequat. Pellentesque posuere orci id
leo molestie, ac rutrum dolor pretium. In orci lectus, sollicitudin nec tellus
non, dictum egestas ante. Ut venenatis libero risus, sed auctor lorem
scelerisque at. Suspendisse porta elementum magna, id ornare erat placerat at.
In tincidunt dolor porta justo rhoncus elementum.

\section{Background}
Nullam viverra turpis nisi, id consequat urna pulvinar eu. Praesent ac pulvinar
nisl. Integer vel est a libero molestie mollis id vel tortor. Morbi ante odio,
cursus et volutpat a, pharetra at erat. Pellentesque habitant morbi tristique
senectus et netus et malesuada fames ac turpis egestas. Curabitur dignissim,
purus at auctor laoreet, magna mi auctor nulla, et mollis eros nisl vel lectus.
Integer bibendum purus ac adipiscing bibendum.

\section{Design and Implementation}
Suspendisse potenti. Donec ornare metus sem, vel semper arcu fermentum in. Nunc
vehicula eros turpis, ut rhoncus nunc sodales pulvinar. Duis sed est interdum,
semper mi a, cursus sapien. Nulla elementum congue purus, eu faucibus nibh
congue et. Ut diam metus, euismod ac quam in, gravida dictum turpis. Vivamus
hendrerit facilisis tempor. Curabitur hendrerit orci ac libero vestibulum, sit
amet dictum enim viverra. Nulla rutrum dolor ligula, nec rutrum massa suscipit
volutpat. Aliquam id mi at velit aliquet imperdiet at nec mi. Integer in massa
nec augue dapibus ullamcorper. Aliquam nec sollicitudin elit.

\section{Discussion and Future Work}
In eu gravida ligula, sit amet tincidunt nisi. Interdum et malesuada fames ac
ante ipsum primis in faucibus.  Vestibulum venenatis non sem vitae hendrerit.
Cras accumsan vel mauris ut vestibulum. Pellentesque auctor quam ut nisl
commodo malesuada. Phasellus adipiscing placerat tortor, in convallis massa
cursus eu. Vestibulum ullamcorper lorem vel nunc dictum, elementum pretium
nulla tincidunt.

\section{Conclusions}

Interdum et malesuada fames ac ante ipsum primis in faucibus. Nulla facilisi.
Etiam viverra condimentum accumsan. Mauris sed dolor sed mi dignissim
vulputate. Duis pharetra purus purus, eu ultricies ligula dignissim vel. Sed
nec mauris a turpis luctus placerat id ut ipsum. Fusce quis tincidunt dolor.
Nunc mattis ornare sem, vel consectetur ipsum iaculis ac. Aenean nulla lacus,
iaculis sed tempor nec, euismod suscipit dui.

% References
\bibliographystyle{abbrv}
\bibliography{paper}

%\balancecolumns
\end{document}

### 2014-07-16 ccsw14/paper.tex
Added to introduction

@@
 In this paper, we propose Data Capsules, a system that is designed to enable
-more secure access to sensitive data while maintaining reasonable guarantees of
+flexible access to sensitive data while maintaining reasonable guarantees of
 data privacy.  Data Capsules uses virtualization to provide remote users with a
@@
 \section{Introduction}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Praesent eget
-tristique dolor, sit amet volutpat lacus. Curabitur nec enim id sapien molestie
-feugiat\cite{clark:pct}. Integer adipiscing sapien urna, rhoncus vehicula elit pellentesque at.
-Aliquam pulvinar feugiat dui quis porta. Cras et erat justo. Phasellus at neque
-sed nisl tincidunt consectetur ut eget ante. Donec porta urna ut bibendum
-cursus. Aenean euismod pellentesque turpis vitae placerat. Nam orci turpis,
-pharetra et lacus sit amet, accumsan semper massa.
 
-%Morbi metus nulla, tristique
-%nec arcu a, ornare fermentum lectus. Curabitur interdum libero non massa
-%laoreet commodo. Morbi porta, risus sit amet rhoncus egestas, felis diam
-%sollicitudin massa, eu iaculis purus sem in turpis. Duis commodo facilisis
-%purus non volutpat. Ut ullamcorper metus in ultrices tempus. Integer nec quam
-%vitae justo dictum vehicula a non augue. Proin sit amet gravida libero.
-%
-Donec ornare iaculis nisl, sit amet luctus metus varius ac. Nam eget tincidunt
-magna. Sed convallis pulvinar eros sed consequat. Pellentesque posuere orci id
-leo molestie, ac rutrum dolor pretium. In orci lectus, sollicitudin nec tellus
-non, dictum egestas ante. Ut venenatis libero risus, sed auctor lorem
-scelerisque at. Suspendisse porta elementum magna, id ornare erat placerat at.
-In tincidunt dolor porta justo rhoncus elementum.
+In the age of the Internet, new data is being created at incredible rates
+\cite{digital-universe}.  Along with these waves of new data, new tools have
+in turn been created to help analyze and gain insight from it. Unfortunately,
+access to data is not always a simple matter.  Some types of data may be
+sensitive, including data protected by copyright, data representing the personal
+information of individuals, as well as any data that may contain secrets that
+its owner does not wish to have revealed, for example corporate or state secrets.
 
-Nullam viverra turpis nisi, id consequat urna pulvinar eu. Praesent ac pulvinar
-nisl. Integer vel est a libero molestie mollis id vel tortor. Morbi ante odio,
-cursus et volutpat a, pharetra at erat. Pellentesque habitant morbi tristique
-senectus et netus et malesuada fames ac turpis egestas. Curabitur dignissim,
-purus at auctor laoreet, magna mi auctor nulla, et mollis eros nisl vel lectus.
-Integer bibendum purus ac adipiscing bibendum.
-
-Donec ornare iaculis nisl, sit amet luctus metus varius ac. Nam eget tincidunt
-magna. Sed convallis pulvinar eros sed consequat. Pellentesque posuere orci id
-leo molestie, ac rutrum dolor pretium. In orci lectus, sollicitudin nec tellus
-non, dictum egestas ante. Ut venenatis libero risus, sed auctor lorem
-scelerisque at. Suspendisse porta elementum magna, id ornare erat placerat at.
-In tincidunt dolor porta justo rhoncus elementum.
-
-Nullam viverra turpis nisi, id consequat urna pulvinar eu. Praesent ac pulvinar
-nisl. Integer vel est a libero molestie mollis id vel tortor. Morbi ante odio,
-cursus et volutpat a, pharetra at erat. Pellentesque habitant morbi tristique
-senectus et netus et malesuada fames ac turpis egestas. Curabitur dignissim,
-purus at auctor laoreet, magna mi auctor nulla, et mollis eros nisl vel lectus.
-Integer bibendum purus ac adipiscing bibendum.
+One good example of a system for access to sensitive data is the data for the
+United States Census.  The federal government established the Census Research
+Data Centers to enable researchers access to unpublished Census data for the
+purpose of conducting research that could provide guidance on public policy.
 
@@
 Etiam viverra condimentum accumsan. Mauris sed dolor sed mi dignissim
-vulputate. Duis pharetra purus purus, eu ultricies ligula dignissim vel. Sed
+vulputate. Duis \cite{clark:pct }pharetra purus purus, eu ultricies ligula dignissim vel. Sed
 nec mauris a turpis luctus placerat id ut ipsum. Fusce quis tincidunt dolor.

### 2014-07-17 ccsw14/paper.tex
Modified abstract

@@
 In this paper, we propose Data Capsules, a system that is designed to enable
-flexible access to sensitive data while maintaining reasonable guarantees of
-data privacy.  Data Capsules uses virtualization to provide remote users with a
-secure, privileged environment in which they can access and analyze sensitive
-data using software of their choice, while minimizing the available channels for
-data leaks, including covert channels.  Our early implementation realizes most
-of this protection, with the exception of the end user's remote desktop
-connection.
+access to sensitive data for analysis by trusted remote users, while
+maintaining reasonable guarantees of data security.  Data Capsules uses
+virtualization to provide remote users with a privileged, but secure
+environment into which they can bring arbitrary, and even potentially
+malicious, software or data in order to analyze sensitive data, while
+minimizing the available channels for data leaks.  Our early implementation
+realizes much of this protection, providing a basic framework for secure analysis of data that addresses many aspects of network, storage, and covert channel security.
 
@@
 Etiam viverra condimentum accumsan. Mauris sed dolor sed mi dignissim
-vulputate. Duis \cite{clark:pct }pharetra purus purus, eu ultricies ligula dignissim vel. Sed
+vulputate. Duis \cite{clark:pct} pharetra purus purus, eu ultricies ligula dignissim vel. Sed
 nec mauris a turpis luctus placerat id ut ipsum. Fusce quis tincidunt dolor.

### 2014-07-18 ccsw14/paper.tex
additions to introduction

@@
 One good example of a system for access to sensitive data is the data for the
-United States Census.  The federal government established the Census Research
-Data Centers to enable researchers access to unpublished Census data for the
-purpose of conducting research that could provide guidance on public policy.
+United States Census.  The federal government established Census Research Data
+Centers (RDCs) to enable researchers access to unpublished Census data for the
+purpose of conducting research that could provide beneficial guidance to public
+policy.  Once granted access, a researcher must visit one of the CRDCs in
+person, and will then be permitted to use preallocated machines with preloaded
+software in order to conduct their research.  Upon completion of their research,
+their results are human-audited to protect confidentiality, and then released.
+While this system provides access to Census data, and seems robust to data
+leaks, it is also highly restrictive, requiring researchers to relocate and use
+foreign hardware to conduct research.
+
+More recently, the Centers for Medicare and Medicaid Services began providing
+Medicare and Medicaid data using the CMS Virtual Research Data Center (VRDC) in
+2013.  VRDC provides a more flexible model by allowing researchers to remotely
+access desktop environments provided by a cloud infrastructure.  This model
+relieves researchers of the need to travel to a physical location, and also
+avoids the security risk of researchers having to receive and safeguard physical
+copies of the data, while they are conducting their research.
+
+However, even this remote access model is still restrictive, in that researchers
+are largely limited to the data analysis tools and software that are made
+available to them in their virtual environments, and any new data or software
+one wishes to upload to a research environment must first undergo a review
+process. To solve this substantial inconvenience, we propose a new system which
+we call Data Capsules.  While, to an end user, a Data Capsules environment will
+minimally differ from the type of environment provided by the CMS Virtual
+Research Data Center, its design grants end users privileged access to install
+software of their choice --- even down to the virtualization-compatible
+operating system of their choice --- without requiring a review process and
+without any compromise to the security of the data being provided.
 

### 2014-07-19 ccsw14/paper.tex
Added to intro

@@
 \title{Securing Remote Access to Sensitive Data}
+
+\author{
+Draft Copy
+}
 
@@
 
+Data Capsules makes use of virtualization as a key element in achieving this
+security and usability goal.  Under the Data Capsules use model, a researcher
+would be allocated a virtual machine (henceforth abbreviated as VM, or guest
+VM), which they could then prepare and maintain with whatever software and
+underlying operating system is best suited to the intended purpose.  Each Data
+Capsules VM is endowed with two modes of operation.  In \emph{maintenance} mode,
+the researcher is endowed with full Internet access, which enables them to
+easily configure and maintain the software environment almost as if they were
+using a their own personal machine; however, in maintenance mode, the researcher
+does not have access to the sensitive data.  In order to gain access to this
+data, the user must switch their VM to the second mode, which we call
+\emph{security} mode.  In security mode, the user's VM state is saved, and
+execution continues on an alternate machine that is firewalled to prevent
+Internet access, but which does have access to the sensitive data.
+
+In this manner, once a user has configured their system and is prepared to
+run their analysis on the data, they simply switch to security mode and start
+the analysis.  Any results obtained from this analysis can then be stored in a
+designated \emph{secure storage volume}, which is only attached to the VM in
+secure mode.  Since this result data is only accessible in secure mode, it
+follows that data can only be released from the virtual environment when in
+secure mode.
+
+When a user wishes to further change their software environment in order to
+continue data analysis, they may then switch back to maintenance mode, after
+which the secure storage volume will detatch, and all VM state since the
+previous switch to secure mode is discarded, with the VM continuing to run from
+the point it left off at when last in maintenance mode.  In this way, every
+trace of the operations performed on the sensitive data is erased and forgotten.
+
+Furthermore, assuming any sensitive data is read-only, and is properly backed
+up, Data Capsules functionality means that there is no need to audit installed
+software for malware, as even infected software would be unable to exfiltrate
+the data from the secure mode environment.
+
+In the following sections, we discuss related work, elaborate on the design and implementation of the Data Capsules system, and discuss the current state of the project and potential directions for future work.
+
 \section{Related Work}

### 2014-07-19 ccsw14/paper.tex
Added threat model + security guarantees section

@@
 
-\title{Securing Remote Access to Sensitive Data}
+\title{Improving the Usability of Remote Access to Sensitive Data}
 
@@
 \section{Related Work}
-Donec ornare iaculis nisl, sit amet luctus metus varius ac. Nam eget tincidunt
-magna. Sed convallis pulvinar eros sed consequat. Pellentesque posuere orci id
-leo molestie, ac rutrum dolor pretium. In orci lectus, sollicitudin nec tellus
-non, dictum egestas ante. Ut venenatis libero risus, sed auctor lorem
-scelerisque at. Suspendisse porta elementum magna, id ornare erat placerat at.
-In tincidunt dolor porta justo rhoncus elementum.
 
-\section{Background}
-Nullam viverra turpis nisi, id consequat urna pulvinar eu. Praesent ac pulvinar
-nisl. Integer vel est a libero molestie mollis id vel tortor. Morbi ante odio,
-cursus et volutpat a, pharetra at erat. Pellentesque habitant morbi tristique
-senectus et netus et malesuada fames ac turpis egestas. Curabitur dignissim,
-purus at auctor laoreet, magna mi auctor nulla, et mollis eros nisl vel lectus.
-Integer bibendum purus ac adipiscing bibendum.
+\section{Design}
 
-\section{Design and Implementation}
+We now discuss the design of the Data Capsules system.  This includes the
+functional requirements for the two modes of operation and a discussion of how
+the system avoids various channels for leaking data.
+
+\subsection{Security Guarantees and Threat Model}
+
+The goal of the Data Capsules system is to provide a usable environment for data
+analysis that also provides a reasonably strong (as we elaborate on below)
+guarantee of data privacy from parties other than the end user to whom a VM
+belongs.  Data Capsules is not designed to provide any protection to the
+integrity of sensitive data.  This is assumed trivial, as the sensitive data is
+presumably unchanging, so that making it read-only and creating appropriate
+backup copies should be a sufficient practice to ensure the data's integrity.
+
+The end user is assumed to be largely trustworthy, since they have presumably
+undergone a vetting process before receiving permission to analyze the sensitive
+data.  However, we do recognize the possibility that others could gain temporary
+access to a VM, for example by using a researcher's machine after they
+momentarily step away.  If the device through which an end user accesses their
+remote VM becomes compromised, this could also open the possibility of long-term
+access to the VM.  While the current implementation does not fully address
+either of these possibilities, we discuss the paths we are investigating in
+Section \ref{sec:Discussion}.
+
+Within a user's VM, we make no assumptions about the software environment and
+treat everything as untrusted, down to the operating system.  Data Capsules is
+designed so that malicious software may succeed in accessing the sensitive data,
+but will be incapable of reliably exfiltrating it from the virtual machine.
+Under this design, we consider the following potential channels for data leaks:
+
+\textbf{Storage Channels} It should not be possible for data that is produced
+as the result of an operation whose input includes sensitive data to be stored
+in such a way that it is accessible outside of secure mode.
+
+\textbf{Network Channels} When a VM has access to sensitive data, e.g. when it
+is in secure mode, it should not be able to access any other machine over the
+network, with the exception of machines that are serving the sensitive data.
+
+\textbf{Virtual Machine Access Channels} End users must access their VMs
+through some network channel that connects a physical device (e.g. the user's
+laptop) to the Data Capsules infrastructure.  This could be a remote terminal
+connection such as SSH, or a remote desktop connection such as VNC or RDP.  In
+both cases, steps must be taken to limit or eliminate channel bandwidth in order
+to protect against the consequences of short and long-term compromises.
+
+\textbf{Intramachine Covert Channels} If there are multiple VMs sharing a
+single physical machine, Data Capsules must ensure that it is not possible for
+one VM running in secure mode (and hence with access to sensitive data) to then
+leak that data to a separate coresident VM running in maintenance mode, which
+could then easily relay that data through any of the aforementioned channels.
+
+\textbf{Intermachine Covert Channels} If there are multiple physical machines
+serving VMs, there may be unexpected covert channels between them that could
+allow similar data leaks from secure to maintenance VMs.  Although such attacks
+are likely somewhat difficult to implement and unlikely to have very high
+bandwidth, steps should be taken to minimize any such known channels to the
+highest degree possible without substantially impacting performance.
+
+The virtual machine infrastructure that underlies the Data Capsules system is
+trusted and assumed to be well implemented.  While such an assumption may be
+somewhat controversial, we note its use in prior work~\cite{capsules}, and also
+note the existence of microkernels such as seL4 and OKL4~\cite{sel4, OKL4},
+which serve as a proof of concept that it is possible to create a lightweight
+hypervisor with a reasonable degree of trustworthiness.  However, the
+development of such a hypervisor lies outside the scope of this paper.
+
+\subsection{Design Overview}
+
+Data Capsules makes heavy use of virtualization.  For each system user, there
+are one or more corresponding VMs.  Multiple VMs per user may be useful in
+situations where there is a large repository of data that is separated into
+multiple corpora each with their own access permissions.  In this case, a user
+could potentially have multiple VMs, each with access to a different corpus.
+However, the handling of such an access permissions system is outside the scope
+of this paper.
+
+The fundamental element that provides the guarantees
+
+\subsection{Maintenance Mode}
+
+\section{Implementation}
 Suspendisse potenti. Donec ornare metus sem, vel semper arcu fermentum in. Nunc
@@
 \section{Discussion and Future Work}
+\label{sec:Discussion}
 In eu gravida ligula, sit amet tincidunt nisi. Interdum et malesuada fames ac

### 2014-07-21 ccsw14/paper.tex
added to related work

@@
 
-\title{Improving the Usability of Remote Access to Sensitive Data}
+\title{Improving Usability of Remote Access to Sensitive Data}
 
 \author{
-Draft Copy
+Draft Copy: \today
 }
@@
 
-One good example of a system for access to sensitive data is the data for the
-United States Census.  The federal government established Census Research Data
-Centers (RDCs) to enable researchers access to unpublished Census data for the
-purpose of conducting research that could provide beneficial guidance to public
-policy.  Once granted access, a researcher must visit one of the CRDCs in
-person, and will then be permitted to use preallocated machines with preloaded
-software in order to conduct their research.  Upon completion of their research,
-their results are human-audited to protect confidentiality, and then released.
-While this system provides access to Census data, and seems robust to data
-leaks, it is also highly restrictive, requiring researchers to relocate and use
-foreign hardware to conduct research.
-
-More recently, the Centers for Medicare and Medicaid Services began providing
-Medicare and Medicaid data using the CMS Virtual Research Data Center (VRDC) in
-2013.  VRDC provides a more flexible model by allowing researchers to remotely
-access desktop environments provided by a cloud infrastructure.  This model
-relieves researchers of the need to travel to a physical location, and also
-avoids the security risk of researchers having to receive and safeguard physical
-copies of the data, while they are conducting their research.
+% Mention state of the art for supervised research
 
@@
 
+\subsection{Real-World Systems}
+
+Systems that enable supervised and audited computational research of sensitive
+data date back at least as far as the 1990s, when a Census Research Data Center
+was opened in Boston.  Since the opening of this initial office, the U.S.
+federal government has established several Research Data Centers (RDCs) across
+the country to enable researchers access to unpublished Census data for the
+purpose of conducting research that could provide beneficial guidance to public
+policy.
+
+Potential researchers must first submit a research proposal demonstrating
+scientific merit and potential benefit to Census Bureau programs.  Once granted
+access, a researcher then visits one of the CRDCs in person, and, after swearing
+to protect data confidentiality, will then be permitted to use preallocated
+machines inside the facility with a preloaded suite of software in order to
+conduct their research.  Upon completion of their research, their results are
+human-audited to protect confidentiality, and then released.
+
+While this system provides access to Census data, and seems robust to data
+leaks, it is also highly restrictive, requiring researchers to relocate and use
+foreign hardware and potentially software as well, to conduct their research.
+
+More recently, the Centers for Medicare and Medicaid Services began providing
+Medicare and Medicaid data using the CMS Virtual Research Data Center (VRDC) in
+2013.  VRDC provides a more flexible model by allowing researchers to remotely
+access desktop environments provided by a cloud infrastructure.  This model
+relieves researchers of the need to travel to a physical location, and also
+avoids the security risk of researchers having to receive and safeguard physical
+copies of the data while they are conducting their research, as was formerly the
+case.
+
+However, this state-of-the-art solution still suffers from significant
+drawbacks, in that researchers are largely limited to using the preallocated
+environments supplied by the VRDCs.  In order to upload software or data not
+already available in a virtual desktop, it must first undergo a manual review
+process.  For large software suites, this can at best add significant
+inconvenience, and at worst expose the system to a malicious attack, as there is
+no method for reliably detecting malicious software.  Data Capsules is designed
+so that no review process is necessary to upload software or data to an
+environment, while providing strong guarantees of data privacy by design.
+
+\subsection{Systems Using Virtualization}
+
+% storage capsules, QubesOS
+
+\subsection{Information Flow Control}
+
+% HiStar
+
 \section{Design}
@@
 
-\subsection{Security Guarantees and Threat Model}
+\subsection{Security Goals and Threat Model}
 

### 2014-07-21 ccsw14/paper.tex
finished related work section

@@
 \usepackage{epstopdf}
+\usepackage{datetime}
+\usdate
+\settimeformat{ampmtime}
 
@@
 \author{
-Draft Copy: \today
+Draft Copy: \today\ \currenttime
 }
@@
 minimizing the available channels for data leaks.  Our early implementation
-realizes much of this protection, providing a basic framework for secure analysis of data that addresses many aspects of network, storage, and covert channel security.
+realizes much of this protection, providing a basic framework for secure
+analysis of data that addresses many aspects of network, storage, and covert
+channel security.
 
@@
 
-% Mention state of the art for supervised research
+Current state-of-the-art systems developed for the purpose of enabling access to
+these types of data use remotely accessed machines, which can then be used by
+researchers without the need for travel to conduct their analysis. But beyond
+this apparent convenience, rules for using such systems are often highly
+restrictive, with researchers limited to using predefined software environments,
+and being subject to strict auditing requirements for all software and data that
+is uploaded to these software environments.
 
-However, even this remote access model is still restrictive, in that researchers
-are largely limited to the data analysis tools and software that are made
-available to them in their virtual environments, and any new data or software
-one wishes to upload to a research environment must first undergo a review
-process. To solve this substantial inconvenience, we propose a new system which
-we call Data Capsules.  While, to an end user, a Data Capsules environment will
+To solve this substantial inconvenience, we propose a new system which we call
+Data Capsules.  While, to an end user, a Data Capsules environment will
 minimally differ from the type of environment provided by the CMS Virtual
@@
 data, the user must switch their VM to the second mode, which we call
-\emph{security} mode.  In security mode, the user's VM state is saved, and
+\emph{secure} mode.  In secure mode, the user's VM state is saved, and
 execution continues on an alternate machine that is firewalled to prevent
@@
 In this manner, once a user has configured their system and is prepared to
-run their analysis on the data, they simply switch to security mode and start
+run their analysis on the data, they simply switch to secure mode and start
 the analysis.  Any results obtained from this analysis can then be stored in a
@@
 
-In the following sections, we discuss related work, elaborate on the design and implementation of the Data Capsules system, and discuss the current state of the project and potential directions for future work.
+In the following sections, we discuss related work, elaborate on the design and
+implementation of the Data Capsules system, and discuss the current state of
+the project and potential directions for future work.
 
 \section{Related Work}
+
+As mentioned, in addition to research systems for securing data, there are
+several real-world systems that address the issue of securing sensitive data.
+In this section, we discuss a selection of both real-world implementations of
+such systems as well as systems proposed by the academic community.
 
@@
 
-% storage capsules, QubesOS
+QubesOS \cite{qubes} is a system for isolating a user's desktop environment into
+secure compartments, called ``Qubes," which are intended to prevent malicious
+code from compromising the entire system by restricting its effect to the
+compartment in which it resides.  Although Qubes is very effective for this 
+purpose, it would not serve as a particularly secure tool for data access, as it
+suffers from the same challenge of how to move data from one compartment, which
+is presumably unrestricted, into another compartment containing the sensitive
+data to be operated on.  Data Capsules offers a clear design for allowing the
+import of data and code into an environment that can still offer secure access
+to sensitive data.
+
+Storage Capsules \cite{capsules-borders} is a virtualization-based system for
+protecting data privacy on individual machines.  Under Storage Capsules, a
+machine runs a virtual machine manager (VMM) system, which in turn runs the
+user's operating system environment within a VM.  On the same machine, there is
+a second VM running an interface that serves as a ``storage capsule" for
+sensitive data.  The user's main VM is endowed with two modes of operation,
+\emph{normal} and \emph{secure} mode, and similar to Data Capsules, normal mode
+grants full network priviliges while secure mode grants access to the sensitive
+data, via communication with the storage capsule VM.
+
+While our work shares the dual-mode user interface of Storage Capsules, the
+differences largely end there.  Our system is designed under a significantly
+different usage model, in which end users are accessing the VM remotely and
+operating on data stored outside of the machine a VM resides on. This different
+environment substantially alters the threat model, and provides a different set
+of channels for potential data leaks.  As a result, Data Capsules offers the
+potential to provide an environment for data access that is faster and more
+robust to data leaks.
 
@@
 
-% HiStar
+Mandatory access control systems, such as SELinux \cite{selinux}, offer the
+ability to place strict controls on how a system is used.  It is conceivable
+that a machine could be configured with a policy that would enable users to
+analyze sensitive data without a high risk of data leakage.  However, such a
+policy would have to be written and maintained by an expert, in order to account
+for each piece of software that is to operate on the data, and the policy would
+likely be so large and complex that it would be hard to avoid at least some
+security errors.  Because of this, it is not a particularly flexible solution,
+and would still require manual action before a user could begin using their own
+custom software in such an environment.
+
+Specialized operating systems that enforce information flow control, such as
+HiStar \cite{histar}, offer low-level operating system primitives that allow
+developers to design programs in such a way that data leaks are not possible.
+As with SELinux, there is still a substantial burden of having to craft policies
+that prevent these malicious data flows.  Additionally, programs must be
+specifically developed for compatibility with HiStar, creating an even greater
+burden on users so long as that system is not widely supported by the software
+community.
 
@@
 \section{Implementation}
+
 Suspendisse potenti. Donec ornare metus sem, vel semper arcu fermentum in. Nunc
@@
 \label{sec:Discussion}
+
 In eu gravida ligula, sit amet tincidunt nisi. Interdum et malesuada fames ac
@@
 Etiam viverra condimentum accumsan. Mauris sed dolor sed mi dignissim
-vulputate. Duis \cite{clark:pct} pharetra purus purus, eu ultricies ligula dignissim vel. Sed
-nec mauris a turpis luctus placerat id ut ipsum. Fusce quis tincidunt dolor.
-Nunc mattis ornare sem, vel consectetur ipsum iaculis ac. Aenean nulla lacus,
-iaculis sed tempor nec, euismod suscipit dui.
+vulputate. Duis \cite{clark:pct} pharetra purus purus, eu ultricies ligula
+dignissim vel. Sed nec mauris a turpis luctus placerat id ut ipsum. Fusce quis
+tincidunt dolor.  Nunc mattis ornare sem, vel consectetur ipsum iaculis ac.
+Aenean nulla lacus, iaculis sed tempor nec, euismod suscipit dui.
 
@@
 
-%\balancecolumns
 \end{document}

### 2014-07-22 webservice/src/main/java/edu/indiana/d2i/sloan/SloanWSApplication.java
clean up

@@ -44,19 +44,7 @@ public class SloanWSApplication extends Application {
 			} catch (ClassNotFoundException e) {
 				throw new RuntimeException(e);
 			} 
-		}
-		
-//		s.add(CreateVM.class);
-//		s.add(QueryVM.class);
-//		s.add(LaunchVM.class);
-//		s.add(DeleteVM.class);
-//		s.add(StopVM.class);
-//		s.add(SwitchVM.class);
-//		s.add(ListImage.class);
-//		
-//		s.add(DownloadResult.class);
-//		s.add(UploadResult.class);
-		
+		}		
 		return s;
 	}
 

### 2014-07-23 ccsw14/paper.tex
finished design section

@@
 The virtual machine infrastructure that underlies the Data Capsules system is
-trusted and assumed to be well implemented.  While such an assumption may be
+trusted and assumed to be well implemented.  While this assumption may be
 somewhat controversial, we note its use in prior work~\cite{capsules}, and also
@@
 which serve as a proof of concept that it is possible to create a lightweight
-hypervisor with a reasonable degree of trustworthiness.  However, the
-development of such a hypervisor lies outside the scope of this paper.
+hypervisor with a reasonable degree of trustworthiness based in formal proofs
+of correctness.  However, the development of such a hypervisor lies outside the
+scope of this paper.
 
@@
 
-The fundamental element that provides the guarantees
+The major element that is used to isolate VMs that are in the middle of
+accessing sensitive data is live migration \cite{migration-clark}.  When a VM
+switches from maintenance mode to secure mode, the underlying virtual machine
+manager migrates that VM to a separate physical machine that has access to the
+sensitive data, but which is configured to lack network access for any purpose
+other than to receive other incoming VM migrations.  In other words, VMs in
+maintenance mode are strictly segregated within separate physical machines from
+VMs running in secure mode.  We can therefore refer to a physical machine
+running only maintenance VMs as a \emph{maintenance machine}, and likewise a
+\emph{secure machine} runs only secure mode instantiations of VMs.
+
+By adopting this segregated design, the threat of data leak via covert channels
+within a single physical machine is eliminated, so long as the trusted computing
+base remains uncompromised.  The remaining risk is that of one VM in secure mode
+leaking data to another VM that is also in secure mode, but there is no reason
+to want to do so, as both VMs would already have access to the entire secure
+repository.  On a machine running VMs in maintenance mode, there is no way to
+access the sensitive data, and hence no data to be leaked.  We stress that this
+design also requires at least two physical machines in order to provide these
+security guarantees.
 
 \subsection{Maintenance Mode}
+
+As the name suggests, maintenance mode is intended to allow a user to maintain
+their VM's software environment, performing actions such as updating software
+and uploading any files needed to conduct their research when in secure mode.
+In accordance with these needs, maintenance mode allows the user full Internet
+access and persistent storage (i.e.  changes to the virtual hard disk will
+persist on reboot) in exchange for a lack of access to the sensitive data and
+the user's persistent secure storage volume.
+
+When a user initiates the switch to secure mode, their system state is migrated
+to a secure machine.  Once this migration has completed, the state of the
+maintenance VM is frozen, and remains in this frozen state until the user exits
+secure mode.  Upon exit from secure mode, the maintenance VM is simply unfrozen,
+and continues processing from the point at which it left off at the time of the
+original switch to secure mode.
+
+\subsection{Secure Mode}
+
+When the user switches to secure mode, the state of their VM, previously in
+maintenance mode, is migrated to a secure machine where it continues execution.
+By switching to secure mode, the user sacrifices network access and virtual hard
+disk storage persistence in exchange for access to the sensitive data and their
+secure storage volume, which is mounted upon entering secure mode.  Any changes
+to the normal storage are stored only in the machine's virtual memory, and
+discarded when the user exits secure mode.  Only changes to the user's secure
+storage volume will persist to the next time the user enters secure mode.
+Indeed, when the user switches back to maintenance mode from secure mode, all
+accumulated state change after the original switch into secure mode is discarded
+from memory, and the maintenance VM simply resumes execution from where it left
+off prior to the switch.
+
+Conceptually, we can also think of this functionality in this way: by using the
+same single virtual machine and transporting it, via migration, to a different
+environment where it is subject to a different set of permissions and
+privileges, we can allow the user to transport a custom software environment
+created in an insecure environment over to a secure environment where it can
+then be used to conduct research, and subsequently be discarded.  By giving them
+a secure storage volume in secure mode, the user can save their intermediate
+results before further iterating their software environment, preserving
+flexibility in the case that they require further enhancements to their
+software environment.
+
+The use of two separate modes for a single virtual machine, as opposed to, say,
+using two separate machines or two separate VMs, allows the user to customize
+the software environment they use when the secure data is accessed.  This would
+not be possible without opening high-risk channels or a burdensome and
+inconvenient auditing process under a two (virtual) machine model.
 

### 2014-07-23 ccsw14/paper.tex
finished implementation section

@@
 \subsection{Security Goals and Threat Model}
+\label{sec:Threat Model}
 
@@
 
-Suspendisse potenti. Donec ornare metus sem, vel semper arcu fermentum in. Nunc
-vehicula eros turpis, ut rhoncus nunc sodales pulvinar. Duis sed est interdum,
-semper mi a, cursus sapien. Nulla elementum congue purus, eu faucibus nibh
-congue et. Ut diam metus, euismod ac quam in, gravida dictum turpis. Vivamus
-hendrerit facilisis tempor. Curabitur hendrerit orci ac libero vestibulum, sit
-amet dictum enim viverra. Nulla rutrum dolor ligula, nec rutrum massa suscipit
-volutpat. Aliquam id mi at velit aliquet imperdiet at nec mi. Integer in massa
-nec augue dapibus ullamcorper. Aliquam nec sollicitudin elit.
+A prototype of the Data Capsules system has been developed using the KVM virtual
+machine manager, which is built in to the Linux operating system, along with the
+user-level counterpart application qemu to achieve virtualization.  A set of
+Bash shell scripts were written to implement the basic operations for managing
+Data Capsules VMs.  Although the current prototype does not fully support the
+migration-based mode change described in this paper, migration-based switching
+has been implemented and demonstrated, and the only remaining work is to fully
+integrate it into the existing code.
+
+End users may use either a convenient web interface, or call the scripts
+directly in order to perform basic operations including creation and deletion of
+VMs, running and stopping VMs, as well as switches between maintenance and
+secure modes.
+
+Although it could be argued that other, better choices for a virtual machine
+manager exist on which to base such a demonstration system, it is the opinion of
+the authors that no currently existing virtual machine manager is sufficiently
+well written and designed so as to serve as an ideal choice, since there is no
+x86-based hypervisor that is formally proven to be properly implemented and
+secure.  The choice of KVM was for its ubiquity as part of the Linux operating
+system along with ease of development.  However, as mentioned in Section
+\ref{sec:Threat Model}, microkernels do exist which serve to demonstrate that
+at least something close to such a robust, well-designed VMM could be made.
+
+\subsection{Network Protections}
+
+The Data Capsules prototype uses iptables rules to implement a simple set of
+firewall rules for VMs.  All VMs on a machine reside on a bridged, private
+network.  In secure mode, additional rules are imposed on the private IP address
+of the VM, restricting all network access except that between the VM and the IP
+address of the one or more servers containing the sensitive data, as written in
+a simple configuration file.
+
+Because limiting access to only a small set of servers eliminates access to
+public DNS servers, for our prototype we configured the \texttt{/etc/hosts} file
+on VM images to allow them to resolve the server domain names when in secure
+mode.  More elegant solutions could be achieved by allowing access to a
+custom purpose-built DNS server, or by using solutions like cloud-init, which is
+available for Linux based virtual machines.
+
+\subsection{Secure Storage Volume}
+
+In order for a user's intermediate research to persist across sessions within
+secure mode, a storage volume is attached upon entering secure mode.  This
+volume is implemented as a disk image file, which is attached as an emulated USB
+thumb drive.  From the user's perspective, this is a convenient access paradigm,
+as one can imagine a virtual USB drive being ``plugged in" to the machine when
+the user switches to secure mode, and then being removed when the user switches
+back.
+
+One currently unresolved issue with this approach, and any using a mounted disk
+volume, is the issue of caching by the operating system of any changes to the
+secure volume.  Since changes typically are not immediately written back to a
+disk volume, this raises the possibility that a user could switch out of secure
+mode before all data is written back, resulting in data loss.  Currently, we
+warn users of the prototype before they switch out of secure mode that they must
+detatch the secure volume or risk data loss.  In the future, we might include
+software on preconfigured VMs that can help users know when it is safe to switch
+out of secure mode.
+
+\subsection{Releasing Research Results}
+
+Since users ultimately require some method for releasing the results of their
+research, qemu was configured to create a channel through the serial port on the
+VM to a UNIX domain socket on the host machine.  A simple daemon was written
+which runs on the secure machine to listen on this socket for any results to be
+released.  It follows an extremely simple protocol that takes the file name,
+file size, and file data and exports them to a database where they can then
+undergo a review process.  We discuss potential approaches for automating this
+review process for candidate release data in Section \ref{sec:Results Audit}
 
@@
 \label{sec:Discussion}
+
+%mode switch seamlessness (across machines)
+
+%secure client access w/TPM
+% -secure client
+% -secure VNC connection
+
+%managing migration speeds for performance & network
+
+%automating part/all of result release audit
+%\label{sec:Results Audit}
+
+%evaluation plans and some preliminary results
 

### 2014-07-24 ccsw14/paper.tex
added to discussion section

@@
 
+\textbf{Legitimate Release Channels} When researchers have finished working with
+sensitive data, they must be given some channel for releasing their final
+results.  Since this channel is purpose-built for releasing data that is a
+reflection of the data being protected, safeguards must be put in place to
+ensure that ``too much" information is not leaked.
+
 \textbf{Virtual Machine Access Channels} End users must access their VMs
@@
 
+Although a basic prototype has been implemented and is fully functional, there
+are still many potential routes for improvement.  We discuss many of these,
+including some we are currently investigating and some that we intend to look
+into in the future.
+
+\subsection{Securing Client Access}
+
+As mentioned in Section \ref{sec:Threat Model}, the channels used to access a VM
+are an important element that must be secured in order to safeguard any private
+data.  In the event of more sophisticated attacks, such as a man-in-the-middle
+attack between the end user and the Data Capsules infrastructure or a compromise
+of the end user's device, it is not hard to conceive of very powerful attacks.
+
+For instance, if an end user's device is compromised, the malicious party could
+then install malicious software on the VM while in maintenance mode that is
+designed to enter secure mode and subsequently encode all of the private data in
+the format of the channel medium (e.g. as an image in the case of remode desktop
+protocols like VNC).  The data could then be decoded by additional software
+running on the compromised client, or even forwarded for decoding on the
+attacker's personal infrastructure.  In the case of a 1024x768 screen resolution
+refreshing at 15 times per second, an attacker could leak over 2 GB of data from
+a protected repository within one minute, or at the very least saturate the
+channel connection with the leaked data.
+
+This example clearly illustrates the imperative need to secure the remote access
+channel from these types of attack.  There is also, however, a limit to what can
+be done, since the user ultimately must have some way of seeing into the VM that
+they are remotely administering.  The first obvious step is to secure the
+client-server connection using SSL with certificate-based authentication.  While
+this would seem trivial to accomplish, there is an unfortunate lack of support
+for such a secure connection in existing widely used VNC clients.  We are
+currently in the process of modifying an existing VNC client to support SSL
+connections with certificates.  Beyond this initial setback, we expect
+implementation to be relatively trivial.
+
+Beyond securing the connection, we are currently investigating methods for
+securing the end client device using the Trusted Platform Module (TPM) on an end
+user's device to remotely attest that the end client is secure.  We envision the
+final implementation will consist of a USB thumb drive containing a barebones
+Linux operating system with an installed VNC client, and will use the TPM's
+features to perform a measured boot of Linux and the VNC client.  Users will
+plug in the USB drive to their device and boot into the secure environment in
+order to access their Data Capsules VMs.  In this way, we hope to greatly raise
+the bar for a malicious attacker who wishes to gain access to sensitive data,
+while preserving a reasonably uncumbersome method for access to the end user.
+
+\subsection{Optimizing Mode Switches}
+
 %mode switch seamlessness (across machines)
-
-%secure client access w/TPM
-% -secure client
-% -secure VNC connection
 
@@
 
+\subsection{Partial Automation of Release Auditing}
+\label{sec:Results Audit}
+
 %automating part/all of result release audit
-%\label{sec:Results Audit}
+
+\subsection{Evaluation}
+
+\subsubsection{User Study}
+
+\subsubsection{Performance}
 
 %evaluation plans and some preliminary results
-
-In eu gravida ligula, sit amet tincidunt nisi. Interdum et malesuada fames ac
-ante ipsum primis in faucibus.  Vestibulum venenatis non sem vitae hendrerit.
-Cras accumsan vel mauris ut vestibulum. Pellentesque auctor quam ut nisl
-commodo malesuada. Phasellus adipiscing placerat tortor, in convallis massa
-cursus eu. Vestibulum ullamcorper lorem vel nunc dictum, elementum pretium
-nulla tincidunt.
 

### 2014-07-25 ccsw14/paper.tex
release auditing

@@
 file size, and file data and exports them to a database where they can then
-undergo a review process.  We discuss potential approaches for automating this
+undergo a review process.  We discuss a potential approach for automating this
 review process for candidate release data in Section \ref{sec:Results Audit}
@@
 
+\subsection{Partial Automation of Release Auditing}
+\label{sec:Results Audit}
+
+Although Data Capsules aims to ease secure access to sensitive data by removing
+the need to audit data and software being imported into a secure environment, it
+does not address how to ease the auditing of results data being released from
+that same environment.  This is a challenging issue, since a careful attacker
+could use any number of methods to disguise the leaking of large amounts of data
+as something that appears entirely innocent.
+
+While there is no evident solution, there are likely many measures that could be
+taken to automate the identification of more obvious leak attempts.  For
+example, simply measuring the information content of a candidate for release,
+say by measuring its Shannon entropy or its compressed file size, if the
+candidate is unusually large then it may be immediately flagged as suspicious.
+This is only the tip of the iceberg, and there are likely more elegant and
+powerful solutions available.  We leave investigation into such a system as
+future work.
+
 \subsection{Optimizing Mode Switches}
@@
 %managing migration speeds for performance & network
-
-\subsection{Partial Automation of Release Auditing}
-\label{sec:Results Audit}
-
-%automating part/all of result release audit
 

### 2014-07-27 ccsw14/paper.tex
finished first draft of paper

@@
 \usepackage{datetime}
+
+% Switch between these two to add/remove todo comments
+\usepackage{color}
+\newcommand{\todo}[2]{{\color{red}\textbf{#2}}}
+%\newcommand{\todo}[2]{#1}
+
 \usdate
@@
 \section{Implementation}
+\label{sec:Implementation}
 
@@
 
-%mode switch seamlessness (across machines)
+As mentioned in Section \ref{sec:Implementation}, the migration-based mode
+switch has currently been implemented as a demonstration, and has yet to be
+fully integrated into the prototype.  We are currently addressing two
+issues that will help usability and performance of this key element.
 
-%managing migration speeds for performance & network
+The first of these issues is the fact that migrating a VM across two physical
+machines currently requires the user to reestablish a remote desktop or shell
+connection with the destination machine.  We intend to set up a proxy connection
+system that will redirect the user's connection automatically, without the need
+for any action on the user end.  Such a proxy would need to reside on a machine
+separate from machines serving VMs, otherwise it would create an intermachine
+covert channel, since there would at some point be a connection link between a
+process on a maintenance machine and a process on a secure machine.
+
+The second issue with designing a mature migration-based mode switch is that of
+speed and performance.  Transferring the state of many machines all running in
+maintenance mode could easily strain even a very robust network.  Therefore it
+is likely that some form of rate limiting will need to be imposed in order to
+relieve network pressure.  The consequence of this is of course that the time it
+takes to switch modes will increase.  We therefore plan to investigate to find
+the best compromise between the two needs of limiting use of network bandwidth
+and maximizing switch performance.
 
@@
 
-\subsubsection{User Study}
+Since the ultimate goal of the Data Capsules system is to improve usability of
+remote access to sensitive data, it is natural that performance is an important
+measure of its success.  Virtual machines must not suffer any slowdown, and the
+new mode switching interface that it introduces must not be inconveniently slow.
+Additionally, direct feedback from actual users of our system can serve as a
+useful measure of its merits in practice.
 
@@
 
-%evaluation plans and some preliminary results
+Fortunately, virtual machine performance is not significantly affected by Data
+Capsules, since it requires no significant changes to qemu or KVM.
+Basic VM management functions also run quickly, with starting and stopping and deleting VMs taking only fractions of a second.
+VM creation has also been made almost instantatneous to the user through clever use of copy-on-write references to base disk images.
+This leaves only the performance of switches between modes.
+
+For our evaluation, we plan to consider two useful measurements for assessing
+the speed of a migration-based mode switch.  Since VM state begins to be
+transferred to the target secure machine immediately upon startup, there is a
+time from startup until the state transfer tapers to a steady stream as most of
+the initial state is completely transferred and only new updates are then
+carried over.  In our tests on two servers linked by a direct wired connection,
+this switch \emph{startup time} took about 20 seconds for a base installation of
+Ubuntu 10.04, which took about 3.5 GB of virtual disk space.  We note that the
+entire virtual hard disk contents are not transferred, but only copy-on-write
+changes to the virtual hard disk, as well as the entire contents of memory and
+CPU registers.  This startup time is easily approximated by the actual execution
+time of a normal, undelayed migration between two machines.
+
+The second element is the actual \emph{execution time} of a switch, assuming
+that the aforementioned startup process has already occurred and there is only a
+small amount of update data that is being periodically generated and
+transferred.  So long as the end user has used a machine for a time greater than
+the startup time, and so long as bandwidth is not constrained so much that the
+new data being generated cannot fully transfer within the window in which it is
+generated, then there should only be a small, negligible delay, that is
+virtually imperceptible to the user.
+
+For our existing prototype, which does not use migration-based mode switches,
+but rather makes use of VM live snapshots, it is not inaccurate to say that its
+functionality is comparable to that of a migration-based mode switch without the
+delay of transferring state across machines.  In this sense, the execution time
+of a snapshot-based mode switch should certainly be no less than the execution
+time of an equivalent migration-based execution time.  In light of this, we can
+use the mode switch performance time for a snapshot-based mode switch to
+approximate the upper bound for switch execution time in the final
+migration-based Data Capsules system.  In practice, we have observed that this
+switch time almost always falls within 5 seconds, and is usually closer to 1 or
+2 seconds.
+
+We also note for clarity that these switch times refer to the switch from
+maintenance mode into secure mode.  Under a migration-based mode switch, the
+switch back to maintenance mode simply requires unpausing a VM and switching
+back the remote desktop or shell connection to the maintenance machine that VM
+is running on.
+
+\subsubsection{User Study}
+
+A user study is another important option for evaluating the undeniably
+unfamiliar paradigm of having two machine modes for a single conceptual machine.
+Although we have no plans to conduct a formal study in the near future, we are
+in the process of implementing a prototype of the system that will be used by
+humanities researchers for secure access to copyrighted materials.  After
+initial alpha testing by a small number of subjects unfamiliar with the system,
+feedback was positive, with no apparent complaints or confusion about how it was
+used.  We leave more formal and in depth user studies for future work.
 
@@
 
-Interdum et malesuada fames ac ante ipsum primis in faucibus. Nulla facilisi.
-Etiam viverra condimentum accumsan. Mauris sed dolor sed mi dignissim
-vulputate. Duis \cite{clark:pct} pharetra purus purus, eu ultricies ligula
-dignissim vel. Sed nec mauris a turpis luctus placerat id ut ipsum. Fusce quis
-tincidunt dolor.  Nunc mattis ornare sem, vel consectetur ipsum iaculis ac.
-Aenean nulla lacus, iaculis sed tempor nec, euismod suscipit dui.
+In this paper, we introduced Data Capsules, a system that enables more usable
+access to sensitive data repositories by eliminating the need to audit data and
+software that are brought in to the secure environment.  Data Capsules presents
+users with a maintenance mode for fully customizing their software environment,
+and allows users to then import this environment into a secured environment
+where they can then perform analysis on sensitive data with limited risk of data
+leak.  Data Capsules achieves this mode paradigm by migrating underlying VMs
+between machines segregated by their access to the network and to the sensitive
+data.  A prototype of the system has been implemented, and initial performance
+data and end user feedback are both promising.
 

### 2014-07-28 ccsw14/paper.tex
fixed bibliography

@@
 
-Current state-of-the-art systems developed for the purpose of enabling access to
-these types of data use remotely accessed machines, which can then be used by
-researchers without the need for travel to conduct their analysis. But beyond
-this apparent convenience, rules for using such systems are often highly
-restrictive, with researchers limited to using predefined software environments,
-and being subject to strict auditing requirements for all software and data that
-is uploaded to these software environments.
+The current state-of-the-art \cite{vrdc} in systems developed for the purpose
+of enabling access to these types of data use remotely accessed machines, which
+can then be used by researchers without the need for travel to conduct their
+analysis. But beyond this apparent convenience, rules for using such systems
+are often highly restrictive, with researchers limited to using predefined
+software environments, and being subject to strict auditing requirements for
+all software and data that is uploaded to these software environments.
 
@@
 was opened in Boston.  Since the opening of this initial office, the U.S.
-federal government has established several Research Data Centers (RDCs) across
-the country to enable researchers access to unpublished Census data for the
-purpose of conducting research that could provide beneficial guidance to public
-policy.
+federal government has established several Research Data Centers (RDCs)
+\cite{census-rdc} across the country to enable researchers access to
+unpublished Census data for the purpose of conducting research that could
+provide beneficial guidance to public policy.
 
@@
 Medicare and Medicaid data using the CMS Virtual Research Data Center (VRDC) in
-2013.  VRDC provides a more flexible model by allowing researchers to remotely
-access desktop environments provided by a cloud infrastructure.  This model
-relieves researchers of the need to travel to a physical location, and also
-avoids the security risk of researchers having to receive and safeguard physical
-copies of the data while they are conducting their research, as was formerly the
-case.
+2013 \cite{vrdc}.  VRDC provides a more flexible model by allowing researchers
+to remotely access desktop environments provided by a cloud infrastructure.
+This model relieves researchers of the need to travel to a physical location,
+and also avoids the security risk of researchers having to receive and
+safeguard physical copies of the data while they are conducting their research,
+as was formerly the case.
 
@@
 trusted and assumed to be well implemented.  While this assumption may be
-somewhat controversial, we note its use in prior work~\cite{capsules}, and also
-note the existence of microkernels such as seL4 and OKL4~\cite{sel4, OKL4},
-which serve as a proof of concept that it is possible to create a lightweight
-hypervisor with a reasonable degree of trustworthiness based in formal proofs
-of correctness.  However, the development of such a hypervisor lies outside the
-scope of this paper.
+somewhat controversial, we note its use in prior work~\cite{capsules-borders},
+and also note the existence of microkernels such as seL4 and OKL4~\cite{sel4,
+sel4-secure, OKL4}, which serve as a proof of concept that it is possible to
+create a lightweight hypervisor with a reasonable degree of trustworthiness
+based in formal proofs of correctness.  However, the development of such a
+hypervisor lies outside the scope of this paper.
 
@@
 Because limiting access to only a small set of servers eliminates access to
-public DNS servers, for our prototype we configured the \texttt{/etc/hosts} file
-on VM images to allow them to resolve the server domain names when in secure
-mode.  More elegant solutions could be achieved by allowing access to a
-custom purpose-built DNS server, or by using solutions like cloud-init, which is
-available for Linux based virtual machines.
+public DNS servers, for our prototype we configured the \texttt{/etc/hosts}
+file on VM images to allow them to resolve the server domain names when in
+secure mode.  More elegant solutions could be achieved by allowing access to a
+custom purpose-built DNS server, or by using solutions like
+cloud-init\cite{cloudinit}, which is available for Ubuntu-based virtual
+machines.
 

### 2014-07-29 webservice/src/main/resources/resultrelease.py
add release script

@@ -1,0 +1,113 @@
+#!/usr/bin/python
+
+import MySQLdb
+import cmd
+import sys
+import datetime  
+import smtplib
+from email.MIMEText import MIMEText
+
+######
+DB_HOST = ""
+DB_USER = ""
+DB_PWD = ""
+DB_NAME = ""
+SMTP_HOST = ""
+SMTP_PORT = 587
+SMTP_USER = ""
+SMTP_PWD = ""
+
+DOWNLOAD_URL = ""
+######
+
+class ResultRelease(cmd.Cmd):
+     
+    def __init__(self):
+        cmd.Cmd.__init__(self)
+        self.prompt = '> '
+         
+        self.db = MySQLdb.connect(DB_HOST, DB_USER, DB_PWD, DB_NAME)
+        self.cursor = self.db.cursor()
+        self.email_content_template = "Dear {}, \n\nThank you for using HTRC Data Capsule! " + \
+            "You can download your result from the link below. \n{}"
+    
+    def _query_sql_(self, sql):        
+        try:
+           self.cursor.execute(sql)
+           results = self.cursor.fetchall()
+           for row in results:
+              vmid = row[0] 
+              rid = row[1]
+              starttime = row[3]
+              print("vmid={}, rid={}, starttime={}".format(vmid, rid, starttime))
+        except:
+           print "Error: unable to fecth data"
+           
+    def _send_email_(self, content, subject, destination):
+        msg = MIMEText(content, 'plain')
+        msg['Subject'] = subject
+        msg['From'] = SMTP_USER
+    
+        conn = smtplib.SMTP(SMTP_HOST, SMTP_PORT)
+        conn.set_debuglevel(False)
+        conn.starttls()
+        conn.login(SMTP_USER, SMTP_PWD)
+        try:
+            conn.sendmail(SMTP_USER, destination, msg.as_string())
+        finally:
+            conn.close()        
+         
+    def do_show_unrelease(self, args):
+        sql = "select * from results where notified=0"
+        self._query_sql_(sql)
+        
+    def do_show_release(self, args):
+        sql = "select * from results where notified=1"
+        self._query_sql_(sql)
+        
+    def do_release(self, args):
+        rid = args
+        
+        # check if it has been released or not
+        
+        # update starttime
+        now = datetime.datetime.now()
+        sql = "update results set starttime='{}' where resultid='{}'". \
+            format(now.strftime('%Y-%m-%d %H:%M:%S'), rid)
+        try:
[…]

### 2014-07-29 ccsw14/paper.tex
added figures and more small fixes

@@
 
+\usepackage{url}
 \usepackage{epstopdf}
@@
 
-Data Capsules makes heavy use of virtualization.  For each system user, there
-are one or more corresponding VMs.  Multiple VMs per user may be useful in
-situations where there is a large repository of data that is separated into
-multiple corpora each with their own access permissions.  In this case, a user
-could potentially have multiple VMs, each with access to a different corpus.
-However, the handling of such an access permissions system is outside the scope
-of this paper.
+\begin{figure}[ht!]
+\center
+\includegraphics[width=3.0in]{figures/architecture}
+
+\caption{A Data Capsules virtual machine runs in one of two modes.  In
+maintenance mode, VM state such as memory, CPU and disk state persists across
+shutdowns and mode transitions.  In secure mode, this virtual state is lost
+once a user returns to maintenance mode. The only exception is any data stored
+in the user's secure storage volume, which is attached to the VM only when the
+user enters secure mode.}
+
+\label{fig:architecture}
+\end{figure}
+
+Data Capsules features an architecture, illustrated in Figure
+\ref{fig:architecture}, that makes heavy use of virtualization.  For each
+system user, there are one or more corresponding VMs.  Multiple VMs per user
+may be useful in situations where there is a large repository of data that is
+separated into multiple corpora each with their own access permissions.  In
+this case, a user could potentially have multiple VMs, each with access to a
+different corpus.  However, the handling of such an access permissions system
+is outside the scope of this paper.
 
@@
 Beyond securing the connection, we are currently investigating methods for
-securing the end client device using the Trusted Platform Module (TPM) on an end
-user's device to remotely attest that the end client is secure.  We envision the
-final implementation will consist of a USB thumb drive containing a barebones
-Linux operating system with an installed VNC client, and will use the TPM's
-features to perform a measured boot of Linux and the VNC client.  Users will
-plug in the USB drive to their device and boot into the secure environment in
-order to access their Data Capsules VMs.  In this way, we hope to greatly raise
-the bar for a malicious attacker who wishes to gain access to sensitive data,
-while preserving a reasonably uncumbersome method for access to the end user.
+securing the end client device using the Trusted Platform Module (TPM)
+\cite{tpm} on an end user's device to remotely attest that the end client is
+secure.  We envision the final implementation will consist of a USB thumb drive
+containing a barebones Linux operating system with an installed VNC client, and
+will use the TPM's features to perform a measured boot of Linux and the VNC
+client.  Users will plug in the USB drive to their device and boot into the
+secure environment in order to access their Data Capsules VMs.  In this way, we
+hope to greatly raise the bar for a malicious attacker who wishes to gain
+access to sensitive data, while preserving a reasonably uncumbersome method for
+access to the end user.
 

### 2014-07-29 ccsw14/paper.tex
added deployment example figure

@@
 running only maintenance VMs as a \emph{maintenance machine}, and likewise a
-\emph{secure machine} runs only secure mode instantiations of VMs.
+\emph{secure machine} runs only secure mode instantiations of VMs.  We illustrate an example deployment in Figure \ref{fig:configuration}.
+
+\begin{figure}[ht!]
+\center
+\includegraphics[width=3.0in]{figures/configuration}
+
+\caption{An example deployment configuration for a Data Capsules system.
+Dotted boxes denote physical separation, which is necessary between maintenance
+machines, secure machines, and the remote access proxy.  The sensitive data
+server may exist separately, or it may reside on one of the secure machines.
+The only connections between maintenance and secure machines will be for the
+purpose of tranferring virtual machine state to secure machines, and for
+read-only accesses to the VM disk images.  This data largely travels in one
+direction, from maintenance machines to secure machines.}
+
+\label{fig:configuration}
+\end{figure}
 

### 2014-07-30 ccsw14/paper.tex
many changes; late draft of paper

@@
 
-\title{Improving Usability of Remote Access to Sensitive Data}
+\title{Data Capsules: A Cloud Service for Secure Remote Access to Sensitive Data}
 
 \author{
+%Submission to CCSW 2014
 Draft Copy: \today\ \currenttime
@@
 %       \affaddr{Computer Science and Engineering Division}\\
-%       \affaddr{University of Mighigan}\\
+%       \affaddr{University of Michigan}\\
 %       \affaddr{Ann Arbor, MI, USA}\\
@@
 
-In this paper, we propose Data Capsules, a system that is designed to enable
+In this paper, we propose Data Capsules, a system that is designed to simplify
 access to sensitive data for analysis by trusted remote users, while
 maintaining reasonable guarantees of data security.  Data Capsules uses
-virtualization to provide remote users with a privileged, but secure
+virtualization to provide remote users with a fully privileged, but secure
 environment into which they can bring arbitrary, and even potentially
@@
 analysis of data that addresses many aspects of network, storage, and covert
-channel security.
+channel security, and has been adopted for use in the HathiTrust Research
+Center.
 
@@
 
-\keywords{data capsules, data privacy}
+\keywords{data capsules, data privacy, virtualization, cloud computing}
 
@@
 In the age of the Internet, new data is being created at incredible rates
-\cite{digital-universe}.  Along with these waves of new data, new tools have
-in turn been created to help analyze and gain insight from it. Unfortunately,
+\cite{digital-universe}.  Along with this wave of new data, new tools have in
+turn been created to help analyze and gain insight from it. Unfortunately,
 access to data is not always a simple matter.  Some types of data may be
-sensitive, including data protected by copyright, data representing the personal
-information of individuals, as well as any data that may contain secrets that
-its owner does not wish to have revealed, for example corporate or state secrets.
+sensitive, including data protected by copyright, data representing the
+personal information of individuals, as well as any data that may contain
+secrets that its owner does not wish to have revealed, for example corporate or
+state secrets.
 
-The current state-of-the-art \cite{vrdc} in systems developed for the purpose
+The current state-of-the-art \cite{vrdc} in systems deployed for the purpose
 of enabling access to these types of data use remotely accessed machines, which
-can then be used by researchers without the need for travel to conduct their
+can be used by researchers without the need for travel to conduct their
 analysis. But beyond this apparent convenience, rules for using such systems
@@
 software environments, and being subject to strict auditing requirements for
-all software and data that is uploaded to these software environments.
+all software and data that is uploaded.
 
@@
 Data Capsules.  While, to an end user, a Data Capsules environment will
-minimally differ from the type of environment provided by the CMS Virtual
-Research Data Center, its design grants end users privileged access to install
-software of their choice --- even down to the virtualization-compatible
+minimally differ from the type of environment provided by the aforementioned
+state-of-the-art system, its design grants end users privileged access to
+install software of their choice --- even down to the virtualization-compatible
 operating system of their choice --- without requiring a review process and
-without any compromise to the security of the data being provided.
+without any added theoretical risks to the security of the data being provided.
+
+Indeed, any system handling sensitive data must be carefully designed to close
+potential channels for leaking that data as thoroughly as possible.  Our plans
+for Data Capsules intend for the system to eventually address threats ranging
+from more aggressive attacks, such as attempts to leak data over the network or
+other conventional channels, to more subtle attacks, such as an attempt by an
+attacker to leak large volumes of sensitive data by surreptitiously exploiting
+the trusted user's own remote connection to the machine.  And our current
+prototype is already capable of defending against many types of attacks, both
+large and small, which we discuss in our description of the system's design.
 
@@
 underlying operating system is best suited to the intended purpose.  Each Data
-Capsules VM is endowed with two modes of operation.  In \emph{maintenance} mode,
-the researcher is endowed with full Internet access, which enables them to
+Capsules VM is endowed with two modes of operation.  In \emph{maintenance}
+mode, the researcher is granted full Internet access, which enables them to
 easily configure and maintain the software environment almost as if they were
-using a their own personal machine; however, in maintenance mode, the researcher
-does not have access to the sensitive data.  In order to gain access to this
-data, the user must switch their VM to the second mode, which we call
+using a their own personal machine; however, in maintenance mode, the
+researcher does not have access to the sensitive data.  In order to gain access
+to this data, the user must switch their VM to the second mode, which we call
 \emph{secure} mode.  In secure mode, the user's VM state is saved, and
@@
 continue data analysis, they may then switch back to maintenance mode, after
-which the secure storage volume will detatch, and all VM state since the
+which the secure storage volume is detached and all VM state since the
 previous switch to secure mode is discarded, with the VM continuing to run from
@@
 
-Furthermore, assuming any sensitive data is read-only, and is properly backed
+Furthermore, assuming all sensitive data is read-only, and is properly backed
 up, Data Capsules functionality means that there is no need to audit installed
@@
 leaks, it is also highly restrictive, requiring researchers to relocate and use
-foreign hardware and potentially software as well, to conduct their research.
+foreign hardware, and potentially software as well, to conduct their research.
 
@@
 process.  For large software suites, this can at best add significant
-inconvenience, and at worst expose the system to a malicious attack, as there is
-no method for reliably detecting malicious software.  Data Capsules is designed
-so that no review process is necessary to upload software or data to an
-environment, while providing strong guarantees of data privacy by design.
+inconvenience, and at worst expose the system to a malicious attack, as there
+is no known method for reliably detecting malicious software.  Data Capsules is
+designed so that no review process is necessary to upload software or data to
+an environment, while providing strong guarantees of data privacy by design.
 
@@
 
-QubesOS \cite{qubes} is a system for isolating a user's desktop environment into
-secure compartments, called ``Qubes," which are intended to prevent malicious
-code from compromising the entire system by restricting its effect to the
-compartment in which it resides.  Although Qubes is very effective for this 
-purpose, it would not serve as a particularly secure tool for data access, as it
-suffers from the same challenge of how to move data from one compartment, which
-is presumably unrestricted, into another compartment containing the sensitive
-data to be operated on.  Data Capsules offers a clear design for allowing the
-import of data and code into an environment that can still offer secure access
-to sensitive data.
+QubesOS \cite{qubes} is a system for isolating a user's desktop environment
+into secure compartments, called ``Qubes," which are intended to prevent
+malicious code from compromising the entire system by restricting its effect to
+the compartment in which it resides.  Although Qubes is very effective for this
+purpose, it would not serve as a particularly effective tool for data access,
+as it suffers from the same challenge of how to move data from one compartment,
+which is presumably unrestricted, into another compartment containing the
+sensitive data to be operated on.  Data Capsules offers a clear design for
+allowing the import of data and code into an environment that can still offer
+secure access to sensitive data.
 
@@
 \emph{normal} and \emph{secure} mode, and similar to Data Capsules, normal mode
-grants full network priviliges while secure mode grants access to the sensitive
+grants full network privileges while secure mode grants access to the sensitive
 data, via communication with the storage capsule VM.
@@
 specifically developed for compatibility with HiStar, creating an even greater
-burden on users so long as that system is not widely supported by the software
+burden on users so long as that system is not widely supported by the developer
 community.
+
+Another potential approach to controlling information flow in databases is that
+of differential privacy \cite{differential, differential-survey}. This uses
+statistical methods to ensure that information about individual entries in a
+database is not revealed, while still allowing statistical analysis of the
+underlying data.  While differential privacy is indeed promising, this approach
+clearly has very limited applications, as it only applies to highly structured
+data with obvious approaches for statistical analysis.  Data Capsules allows
+analysis of all types of data, structured or unstructured, with all types of
+software, while maintaining reasonable guarantees of security.
 
@@
 
-The virtual machine infrastructure that underlies the Data Capsules system is
-trusted and assumed to be well implemented.  While this assumption may be
-somewhat controversial, we note its use in prior work~\cite{capsules-borders},
-and also note the existence of microkernels such as seL4 and OKL4~\cite{sel4,
-sel4-secure, OKL4}, which serve as a proof of concept that it is possible to
-create a lightweight hypervisor with a reasonable degree of trustworthiness
-based in formal proofs of correctness.  However, the development of such a
-hypervisor lies outside the scope of this paper.
-
-\subsection{Design Overview}
-
 \begin{figure}[ht!]
@@
 \end{figure}
+
+The virtual machine infrastructure that underlies the Data Capsules system is
+trusted and assumed to be well implemented.  While this assumption may be
+somewhat controversial, we note its use in prior work~\cite{capsules-borders},
+and also note the existence of microkernels such as seL4 and OKL4~\cite{sel4,
+sel4-secure, OKL4}, which serve as a proof of concept that it is possible to
+create a lightweight hypervisor with a reasonable degree of trustworthiness
+based in formal proofs of correctness.  However, the development of such a
+hypervisor lies outside the scope of this paper.
+
+The choice of virtualization versus more lightweight containment approaches,
+such as Linux containers \cite{lxc}, is based both on this greater
+trustworthiness of a smaller hypervisor versus having to trust an entire
+operating system kernel, as well as the benefit of users having greater freedom
+to choose their own virtualization-compatible operating system, rather than
+having a single software environment chosen for them.
+
+\subsection{Design Overview}
 
@@
 sensitive data, but which is configured to lack network access for any purpose
-other than to receive other incoming VM migrations.  In other words, VMs in
-maintenance mode are strictly segregated within separate physical machines from
-VMs running in secure mode.  We can therefore refer to a physical machine
+other than to receive other incoming VM migrations and access read-only VM disk
+images that are served over a network-based file system.  In other words, VMs
+in maintenance mode are strictly segregated within separate physical machines
+from VMs running in secure mode.  We can therefore refer to a physical machine
 running only maintenance VMs as a \emph{maintenance machine}, and likewise a
-\emph{secure machine} runs only secure mode instantiations of VMs.  We illustrate an example deployment in Figure \ref{fig:configuration}.
+\emph{secure machine} runs only secure mode instantiations of VMs.  We
+illustrate an example deployment in Figure \ref{fig:configuration}.
+
+By adopting this segregated design, the threat of data leak via covert channels
+within a single physical machine is eliminated, so long as the trusted
+computing base remains uncompromised.  Intermachine covert channels along the
+connection between maintenance machines and secure machines are the greatest
+threat.  However, this threat is relatively low, since data largely travels one
+way from maintenance machines to secure machines.  We plan to consider more
+robust, active protections to this channel as future work.  The remaining risk
+is that of one VM in secure mode leaking data to another VM that is also in
+secure mode, but there is no reason to want to do so, as both VMs would already
+have access to the entire secure repository.  On a machine running VMs in
+maintenance mode, there is no way to access the sensitive data, and hence no
+data to be leaked.  We stress that this design also requires at least two
+physical machines in order to provide these security guarantees.
 
@@
 The only connections between maintenance and secure machines will be for the
-purpose of tranferring virtual machine state to secure machines, and for
+purpose of transferring virtual machine state to secure machines, and for
 read-only accesses to the VM disk images.  This data largely travels in one
@@
 \end{figure}
-
-By adopting this segregated design, the threat of data leak via covert channels
-within a single physical machine is eliminated, so long as the trusted computing
-base remains uncompromised.  The remaining risk is that of one VM in secure mode
-leaking data to another VM that is also in secure mode, but there is no reason
-to want to do so, as both VMs would already have access to the entire secure
-repository.  On a machine running VMs in maintenance mode, there is no way to
-access the sensitive data, and hence no data to be leaked.  We stress that this
-design also requires at least two physical machines in order to provide these
-security guarantees.
 
@@
 access and persistent storage (i.e.  changes to the virtual hard disk will
-persist on reboot) in exchange for a lack of access to the sensitive data and
-the user's persistent secure storage volume.
+persist across reboots and mode changes) in exchange for a lack of access to
+the sensitive data and the user's persistent secure storage volume.
 
-When a user initiates the switch to secure mode, their system state is migrated
-to a secure machine.  Once this migration has completed, the state of the
-maintenance VM is frozen, and remains in this frozen state until the user exits
-secure mode.  Upon exit from secure mode, the maintenance VM is simply unfrozen,
-and continues processing from the point at which it left off at the time of the
+As a user works in maintenance mode, the system state of their VM is
+synchronized to a secure machine in anticipation of an eventual switch to
+secure mode.  When a user initiates the switch to secure mode, the migration of
+system state to the secure machine is completed, the state of the maintenance
+VM is frozen, and it remains in this frozen state until the user exits secure
+mode.  Upon exit from secure mode, the maintenance VM is simply unfrozen, and
+continues processing from the point at which it left off at the time of the
 original switch to secure mode.
@@
 maintenance mode, is migrated to a secure machine where it continues execution.
-By switching to secure mode, the user sacrifices network access and virtual hard
-disk storage persistence in exchange for access to the sensitive data and their
-secure storage volume, which is mounted upon entering secure mode.  Any changes
-to the normal storage are stored only in the machine's virtual memory, and
-discarded when the user exits secure mode.  Only changes to the user's secure
-storage volume will persist to the next time the user enters secure mode.
-Indeed, when the user switches back to maintenance mode from secure mode, all
-accumulated state change after the original switch into secure mode is discarded
-from memory, and the maintenance VM simply resumes execution from where it left
-off prior to the switch.
+By switching to secure mode, the user sacrifices network access and virtual
+hard disk storage persistence in exchange for access to the sensitive data and
+their secure storage volume, which is mounted upon entering secure mode.  Any
+changes to the normal storage are stored only in the secure machine's virtual
+memory, and discarded when the user exits secure mode.  Only changes to the
+user's secure storage volume will persist to the next time the user enters
+secure mode.  Indeed, when the user switches back to maintenance mode from
+secure mode, all accumulated state change after the original switch into secure
+mode is discarded from memory, and the maintenance VM simply resumes execution
+from where it left off prior to the switch.
 
@@
 
-A prototype of the Data Capsules system has been developed using the KVM virtual
-machine manager, which is built in to the Linux operating system, along with the
-user-level counterpart application qemu to achieve virtualization.  A set of
-Bash shell scripts were written to implement the basic operations for managing
-Data Capsules VMs.  Although the current prototype does not fully support the
-migration-based mode change described in this paper, migration-based switching
-has been implemented and demonstrated, and the only remaining work is to fully
-integrate it into the existing code.
+A prototype of the Data Capsules system has been developed using the KVM
+virtual machine hypervisor, which is built in to the Linux operating system,
+along with the user-level counterpart application qemu to achieve
+virtualization.  A set of Bash shell scripts were written to implement the
+basic operations for managing Data Capsules VMs.  Although the current
+prototype does not fully support the migration-based mode change described in
+this paper, migration-based switching has been implemented and demonstrated,
+and the only remaining work is to fully integrate it into the existing code.
 
@@
 firewall rules for VMs.  All VMs on a machine reside on a bridged, private
-network.  In secure mode, additional rules are imposed on the private IP address
-of the VM, restricting all network access except that between the VM and the IP
-address of the one or more servers containing the sensitive data, as written in
-a simple configuration file.
+network.  In secure mode, additional rules are imposed on the private IP
+address(es) of the VM, restricting all network access except that between the
+VM and the IP address of the one or more servers containing the sensitive data,
+as written in a simple configuration file.
 
@@
 mode before all data is written back, resulting in data loss.  Currently, we
-warn users of the prototype before they switch out of secure mode that they must
-detatch the secure volume or risk data loss.  In the future, we might include
-software on preconfigured VMs that can help users know when it is safe to switch
-out of secure mode.
+warn users of the prototype before they switch out of secure mode that they
+must first detach the secure volume or risk data loss.  In the future, we might
+include software on popular preconfigured VMs that can help users know when it
+is safe to switch out of secure mode.
 
@@
 Since users ultimately require some method for releasing the results of their
-research, qemu was configured to create a channel through the serial port on the
-VM to a UNIX domain socket on the host machine.  A simple daemon was written
-which runs on the secure machine to listen on this socket for any results to be
-released.  It follows an extremely simple protocol that takes the file name,
-file size, and file data and exports them to a database where they can then
-undergo a review process.  We discuss a potential approach for automating this
-review process for candidate release data in Section \ref{sec:Results Audit}
+research, qemu was configured to create a channel through the serial port on
+the VM to a UNIX domain socket on the host machine.  A simple daemon was
+written which runs on the secure machine to listen on this socket for any
+results to be released.  It follows an extremely simple protocol that takes the
+file name, file size, and file data and exports them to a database where they
+can then undergo a review process.  We discuss a potential approach for
+partially automating this review process for candidate release data in Section
+\ref{sec:Results Audit}.
 
@@
 Although a basic prototype has been implemented and is fully functional, there
-are still many potential routes for improvement.  We discuss many of these,
-including some we are currently investigating and some that we intend to look
-into in the future.
+are still many potential routes for improvement.  We discuss several of these,
+including some that we are currently investigating and some that we intend to
+look into in the future.
 
@@
 designed to enter secure mode and subsequently encode all of the private data in
-the format of the channel medium (e.g. as an image in the case of remode desktop
+the format of the channel medium (e.g. as an image in the case of remote desktop
 protocols like VNC).  The data could then be decoded by additional software
@@
 
-This example clearly illustrates the imperative need to secure the remote access
-channel from these types of attack.  There is also, however, a limit to what can
-be done, since the user ultimately must have some way of seeing into the VM that
-they are remotely administering.  The first obvious step is to secure the
-client-server connection using SSL with certificate-based authentication.  While
-this would seem trivial to accomplish, there is an unfortunate lack of support
-for such a secure connection in existing widely used VNC clients.  We are
-currently in the process of modifying an existing VNC client to support SSL
-connections with certificates.  Beyond this initial setback, we expect
-implementation to be relatively trivial.
+This example clearly illustrates the imperative need to secure the remote
+access channel from these types of attacks.  There is also, however, a limit to
+what can be done, since the user ultimately must have some way of seeing into
+the VM that they are remotely administering.  The first obvious step is to
+secure the client-server connection using SSL with certificate-based
+authentication.  While this would seem trivial to accomplish, there is an
+unfortunate lack of support for such a secure connection in existing widely
+used VNC clients.  We are currently in the process of modifying an existing VNC
+client to support SSL connections with certificates.  Beyond this initial
+setback, we expect implementation to be relatively trivial.
 
@@
 \cite{tpm} on an end user's device to remotely attest that the end client is
-secure.  We envision the final implementation will consist of a USB thumb drive
-containing a barebones Linux operating system with an installed VNC client, and
-will use the TPM's features to perform a measured boot of Linux and the VNC
-client.  Users will plug in the USB drive to their device and boot into the
+secure.  We envision the final implementation consisting of a USB thumb drive
+containing a barebones Linux operating system with an installed VNC client,
+which will use the TPM's features to perform a measured boot of Linux and the
+VNC client.  Users will plug in the USB drive to their device and boot into the
 secure environment in order to access their Data Capsules VMs.  In this way, we
 hope to greatly raise the bar for a malicious attacker who wishes to gain
-access to sensitive data, while preserving a reasonably uncumbersome method for
-access to the end user.
+access to sensitive data by attacking the remote access connection, while
+preserving a reasonably uncumbersome method for access to the end user.
 
@@
 Although Data Capsules aims to ease secure access to sensitive data by removing
-the need to audit data and software being imported into a secure environment, it
-does not address how to ease the auditing of results data being released from
-that same environment.  This is a challenging issue, since a careful attacker
-could use any number of methods to disguise the leaking of large amounts of data
-as something that appears entirely innocent.
+the need to audit data and software being imported into a secure environment,
+it does not address how to ease the auditing of results data being released
+from that same environment.  This is a challenging issue, since a careful
+attacker could potentially use any number of methods to disguise the leaking of
+large amounts of data as something that appears entirely innocent.
 
@@
 The second issue with designing a mature migration-based mode switch is that of
-speed and performance.  Transferring the state of many machines all running in
-maintenance mode could easily strain even a very robust network.  Therefore it
-is likely that some form of rate limiting will need to be imposed in order to
-relieve network pressure.  The consequence of this is of course that the time it
-takes to switch modes will increase.  We therefore plan to investigate to find
-the best compromise between the two needs of limiting use of network bandwidth
-and maximizing switch performance.
+speed and performance.  Transferring the state of many machines all
+simultaneously running in maintenance mode could easily strain even a very
+robust network.  Therefore, it is likely that some form of rate limiting will
+need to be imposed in order to relieve network pressure.  The consequence of
+this is of course that the time it takes to switch modes will increase.  We
+therefore plan to investigate to find the best compromise between the two needs
+of limiting use of network bandwidth and maximizing switch performance.
+
+\subsection{Utilizing Excess Computational Power}
+
+One important design choice made in the Data Capsules system is the segregation
+of VMs running in maintenance mode from those running in secure mode.  This
+choice adds robustness to some types of threats, but in exchange requires
+double the computational power to realize an actual deployment.  Since VMs can
+only exist in at most one mode at a time, there is a high potential for wasted
+computational power in practice.  While this is not as much of an issue for
+maintenance machines, which face no significant restrictions, besides not having
+access to the sensitive data, and hence could put their resources to any
+desired use, secure machines are much more complicated.
+
+While we leave the investigation into how secured computational resources could
+be used most efficiently as future work, we do stress that the cost of
+deploying a Data Capsules-enabled system is not strictly double the cost of an
+equivalent non-segregated system, since computational resources on maintenance
+machines can be used to offset these additional costs without increased
+security risk.
 
@@
 
-Since the ultimate goal of the Data Capsules system is to improve usability of
+Since the ultimate goal of the Data Capsules system is to simplify usability of
 remote access to sensitive data, it is natural that performance is an important
 measure of its success.  Virtual machines must not suffer any slowdown, and the
-new mode switching interface that it introduces must not be inconveniently slow.
+mode switching interface that it introduces must not be inconveniently slow.
 Additionally, direct feedback from actual users of our system can serve as a
@@
 Fortunately, virtual machine performance is not significantly affected by Data
-Capsules, since it requires no significant changes to qemu or KVM.
-Basic VM management functions also run quickly, with starting and stopping and deleting VMs taking only fractions of a second.
-VM creation has also been made almost instantatneous to the user through clever use of copy-on-write references to base disk images.
-This leaves only the performance of switches between modes.
+Capsules, since it requires no significant changes to qemu or KVM.  Basic VM
+management functions also run quickly, with starting and stopping and deleting
+VMs taking only fractions of a second.  VM creation has also been made almost
+instantaneous to the user through clever use of copy-on-write references to base
+disk images.  This leaves the functional performance of switches between modes.
 
@@
 
+One final issue that requires investigation is memory use in secure mode.  Since
+system state in this mode is not persistent, changes to it are stored in virtual
+memory.  For a more comprehensive evaluation, we intend to investigate this
+memory usage, and if necessary find ways to reduce it or any effects it may have
+on long term performance.
+
 \subsubsection{User Study}
@@
 A user study is another important option for evaluating the undeniably
-unfamiliar paradigm of having two machine modes for a single conceptual machine.
-Although we have no plans to conduct a formal study in the near future, we are
-in the process of implementing a prototype of the system that will be used by
-humanities researchers for secure access to copyrighted materials.  After
-initial alpha testing by a small number of subjects unfamiliar with the system,
-feedback was positive, with no apparent complaints or confusion about how it was
-used.  We leave more formal and in depth user studies for future work.
+unfamiliar paradigm of having two machine modes for a single conceptual
+machine.  Although we have no plans to conduct a formal study in the near
+future, we are in the process of implementing a prototype of the system that
+will be used by humanities researchers for secure access to copyrighted
+material in the HathiTrust Research Center.  After initial alpha testing by a
+small number of subjects unfamiliar with the system, feedback was positive,
+with no apparent complaints or confusion about how it was used.  We leave more
+formal and in depth user studies for future work.
 
@@
 
-In this paper, we introduced Data Capsules, a system that enables more usable
-access to sensitive data repositories by eliminating the need to audit data and
-software that are brought in to the secure environment.  Data Capsules presents
-users with a maintenance mode for fully customizing their software environment,
-and allows users to then import this environment into a secured environment
-where they can then perform analysis on sensitive data with limited risk of data
-leak.  Data Capsules achieves this mode paradigm by migrating underlying VMs
-between machines segregated by their access to the network and to the sensitive
-data.  A prototype of the system has been implemented, and initial performance
-data and end user feedback are both promising.
+In this paper, we introduced Data Capsules, a system that enables simplified
+end-user access to sensitive data repositories by eliminating the need to audit
+data and software that are brought in to the secure environment.  Data Capsules
+presents users with a maintenance mode for fully customizing their software
+environment, and allows users to then import this environment into a secured
+domain where they can then perform analysis on sensitive data with limited risk
+of data leak.  Data Capsules achieves this mode paradigm by migrating
+underlying VMs between machines segregated by their access to the network and
+to the sensitive data.  A prototype of the system has been implemented with
+limited deployment to test users, and initial performance data and end user
+feedback are both promising.
 
